### 2015-12-22 experiment.tex

\documentclass[a4paper,11pt]{article}

\usepackage{fullpage}
\usepackage{courier}
\usepackage{tikz}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{amsmath, amssymb, amsfonts, bm}
\usepackage{algorithm,algorithmic}
\usepackage{amsthm}
\usepackage{float}
\usepackage{graphicx}
\usepackage{booktabs}

\newcommand{\RR}{\mathbb{R}}
\newcommand{\cS}{\mathcal{S}}

\begin{document}

C3UCB Draft \hfill Fall 2015

\vspace*{\baselineskip}

\begin{center}
  \textbf{Contextual Combinatorial Cascading Bandit Experiment}
\end{center}

\vspace*{2\baselineskip}

\section{Preliminary}

\subsection{Synthetic dataset}

Let $\cS = \{x \in \RR^d : \|x\|_2 = 1 \}$ be the unit ball of $\RR^d$. 
Let $E = \{1,...,L\}$ be the set of all base arms. 
We randomly choose $\theta_{\ast}$ with $\|\theta_{\ast}\|_2 = 1$ and randomly assign a $\mu_i \in \cS$ to $i$ for any $i \in E$.
At each time $t$, we choose $b_{t,i} \in \cS$ randomly for any base arm $i$.
Also we fix a constant $h$ to balance weights of $\mu_i$ and disturbance $b_{t,i}$.
Let $x_{t,i} = \frac{\mu_i + h b_{t,i}}{\|\mu_i+h\cdot b_{t,i}\|_2}$ be the context of base arm $i$ at time $t$.
And the weight for base arm $i$ at time $t$ is $w_t(i) = \theta_{\ast}^{\top}x_{t,i} + \epsilon_{t,i}$ where $\epsilon_{t,i} \sim N(\mu_i, \sigma_i)$ for fixed $\sigma_i$.

\subsection{MovieLens}

Let $L$ be the number of all movies and let $M$ be the number of all users. 
The MovieLens dataset is a big matrix $A \in \RR^{M \times L}$ where $A(i,j) \in \{0,1\}$ denotes whether user $i$ has watched movie $j$ or not.
We split $A$ to be $H + F$ by putting entry-$1$ of $A$ to $H$ and $F$ with probability $\sim \mathrm{Ber}(p)$ for some fixed $p$.
We can regard $H$ as know information about history 'What users have watched' and regard $F$ as future criterion.
We use $H$ to derive feature vectors of both users and movies by SVD decomposition $H = USV^{\top}$ where $U = (u_1; ...;u_M)$ and $V = (v_1;...;v_L)$.
At every time $t$, use $x_{t,i} = u_{i}v_j^{\top}$ as the context information of base arm $i$ and randomly choose a user $I_t$.
And use $w_t(j) = F(I_t,j)$ as the weight of base arm $j$.

Notice that for this case, fixed number of base arms, it might have problem if we use $(u_{I_t}, v_j)$ as context information.
Since to find the best arm, it is equivalent to find the best one with highest weights sum, so is equivalent to the best one with highest $\theta_{v}^{\top}x$.

The measurement for MovieLens is accuracy because we don't know the true $\theta_{\ast}$.

\section{Disjunctive case}

\subsection{Need to involve Contextual information}

We experiment both on synthetic data and MovieLens. We compare our method with $\gamma_k = 1$ to the algorithm in Cascading Bandits(ICML'2015) with $L= , K= ,$

\subsection{Need to involve position discount parameter $\gamma$}

We experiment both on synthetic data and MovieLens. We compare our method with $\gamma_k = \gamma^{k-1}$ to the algorithm in Cascading Bandits(ICML'2015) with $L= , K= , \gamma_k=1$.

\subsection{Cascading Information}

We experiment both on synthetic data and MovieLens. We compare our method with $\gamma_k = \gamma^{k-1}$ to the algorithm in Qin Lijing(2014) with $L= , K= , \gamma_k=1$.

\section{Conjunctive case}

\subsection{Need to involve Contextual information}

We experiment on synthetic data. We compare our method with $\gamma_k = 1$ to the algorithm in Cascading Bandits(ICML'2015) with $L= , K= ,$

\subsection{Need to involve position discount parameter $\gamma$}

We experiment on synthetic data. We compare our method with $\gamma_k = \gamma^{k-1}$ to the algorithm in Cascading Bandits(ICML'2015) with $L= , K= , \gamma_k=1$.

\subsection{Cascading Information}

We experiment on synthetic data. We compare our method with $\gamma_k = \gamma^{k-1}$ to the algorithm in Qin Lijing(2014) with $L= , K= , \gamma_k=1$.

\section{Results}

An example output, with $T=10000$, is listed below.

\begin{table}
    \centering
    \renewcommand{\arraystretch}{1.2}
	\begin{tabular}{lc}
	    \toprule
		\textbf{Algorithm}  &\textbf{Cumulative Reward $\sum_{t=1}^Tr_i $}\\
		\midrule
		Li		    &4342.73 \\
		Qin		    &4323.26 \\
		Monkey		    &1765.41 \\
		Kveton		    &1787.81 \\
		Perfect Play		    &N/A \\
		\bottomrule
	\end{tabular}
	\caption{\textbf{Cumulative reward w.r.t different baselines, under Movielens setting.}}
	\label{daily}
\end{table}
\end{document}