### 2016-01-31 contextual combinatorial cascading bandits.tex
update discussion and conclusion

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%% ICML 2016 EXAMPLE LATEX SUBMISSION FILE %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Use the following line _only_ if you're still using LaTeX 2.09.
%\documentstyle[icml2015,epsf,natbib]{article}
% If you rely on Latex2e packages, like most moden people use this:
\documentclass{article}

% use Times
\usepackage{times}
% For figures
\usepackage{graphicx} % more modern
%\usepackage{epsfig} % less modern
\usepackage{subfigure}

% For citations
\usepackage{natbib}

% For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}

% As of 2011, we use the hyperref package to produce hyperlinks in the
% resulting PDF.  If this breaks your system, please commend out the
% following usepackage line and replace \usepackage{icml2016} with
% \usepackage[nohyperref]{icml2016} above.
\usepackage{hyperref}
\hypersetup{
	colorlinks   = true, %Colours links instead of ugly boxes
	urlcolor     = blue, %Colour for external hyperlinks
	linkcolor    = blue, %Colour of internal links
	citecolor   = blue %Colour of citations
}

% Packages hyperref and algorithmic misbehave sometimes.  We can fix
% this with the following command.
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% my settings
\usepackage{amsmath}\allowdisplaybreaks
\usepackage{amssymb,amsfonts,bbm}
\usepackage{mathrsfs}
\usepackage[all]{xy}
\usepackage{empheq}
\usepackage{algorithm,algorithmic}
\usepackage{multirow}
\usepackage{ifthen}
\usepackage[usenames]{color}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{epstopdf}
%\usepackage{epsfig}
%\usepackage{times}
%\usepackage{helvet}
%\usepackage{courier}
%\usepackage{graphicx,subfigure}
%\usepackage{pdfpages}
%\usepackage{multirow}
%\usepackage{multicol}
\newcommand{\bbeta}{\boldsymbol{\beta}}
\newcommand{\EE}{\mathbb{E}}
\newcommand{\II}{\mathbb{I}}
\newcommand{\PP}{\mathbb{P}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\bT}{\mathbb{T}}
\newcommand{\bOne}{\mathbbm{1}}
\newcommand{\bA}{\mathbf{A}}
\newcommand{\ba}{\mathbf{a}}
\newcommand{\bB}{\mathbf{B}}
\newcommand{\bO}{\mathbf{O}}
\newcommand{\br}{\mathbf{r}}
\newcommand{\bU}{\mathbf{U}}
\newcommand{\bV}{\mathbf{V}}
\newcommand{\bw}{\mathbf{w}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\bY}{\mathbf{Y}}
\newcommand{\cA}{\mathcal{A}}
\newcommand{\cB}{\mathcal{B}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cO}{\mathcal{O}}
\newcommand{\cS}{\mathcal{S}}
\newcommand{\argmax}{\mathrm{argmax}}
\newcommand{\trace}{\mathrm{trace}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\| #1 \|}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{remark}[theorem]{Remark}
\newtheorem{proposition}[theorem]{Proposition}%[chapter]
\newtheorem{definition}[theorem]{Definition}%[chapter]
\newtheorem{corollary}[theorem]{Corollary}%[chapter]
\newtheorem{lemma}[theorem]{Lemma}%[chapter]
\newtheorem{example}[theorem]{Example}
\newenvironment{proof}{\noindent {\textbf{Proof. }}}{$\Box$ \medskip}

\newcommand{\CLemmaPrefixExi}{
	Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_K \geq 0$. Let $A = (a_1, ..., a_{\abs{A}})$. For the time $t$ and the conjunctive objective, there exists a prefix $B$ of $A$ such that 
	$$
	p_{t, B} \geq \frac{1}{2}f_{t}^{\ast} - 1 + \gamma_{\abs{B}}, \qquad R(t, B) \geq \frac{1}{2} R(t, A).
	$$ 
}
\newcommand{\CEqDeltaEstAnd}{
	$$
	\EE [R(t, \bA_t) |\cH_t ] \leq \frac{8}{f^{\ast}} \EE \left[ \bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} | \cH_t\right].
	$$
}
\newcommand{\CLemmaSumXiEstimateInDet}{
  If $\lambda \geq C_\gamma$, then
  $$
    \sum_{s=1}^t \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \leq 2\ln \left(\frac{\det(\bV_t)}{\lambda^d} \right).
  $$
}
\newcommand{\CLemmaDetVt}{
  $\det(\bV_t)$ is increasing with respect to $t$ and 
  $$
    \det(\bV_t) \leq (\lambda + C_\gamma t/d)^d.
  $$
}

\newcommand{\hidetext}[1]{}

\newcommand{\compilehidecomments}{false}

% Macro for comments:
\ifthenelse{ \equal{\compilehidecomments}{true} }{%
	\newcommand{\wei}[1]{}
	\newcommand{\yang}[1]{}
	\newcommand{\yajun}[1]{}
}{
\newcommand{\wei}[1]{{\color{blue!50!black}  [\text{Wei:} #1]}}
\newcommand{\shuai}[1]{{\color{brown!60!black} [\text{Shuai:} #1]}}
\newcommand{\shengyu}[1]{{\color{green!50!black} [\text{Shengyu:} #1]}}
}

% Employ the following version of the ``usepackage'' statement for
% submitting the draft version of the paper for review.  This will set
% the note in the first column to ``Under review.  Do not distribute.''
\usepackage{icml2016}

% Employ this version of the ``usepackage'' statement after the paper has
% been accepted, when creating the final version.  This will set the
% note in the first column to ``Proceedings of the...''
%\usepackage[accepted]{icml2016}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Contextual Combinatorial Cascading Bandits}

\begin{document}

\twocolumn[
\icmltitle{Contextual Combinatorial Cascading Bandits}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2015
% package.
\icmlauthor{Your Name}{email@yourdomain.edu}
\icmladdress{Your Fantastic Institute, 314159 Pi St., Palo Alto, CA 94306 USA}
\icmlauthor{Your CoAuthor's Name}{email@coauthordomain.edu}
\icmladdress{Their Fantastic Institute, 27182 Exp St., Toronto, ON M6H 2T1 CANADA}

% You may provide any keywords that you 
% find helpful for describing your paper; these are used to populate 
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{boring formatting information, machine learning, ICML}

\vskip 0.3in
]

\begin{abstract}

We propose the {\it contextual combinatorial cascading bandits}, an online partial monitoring game where at each time step a learning agent is given a set of context information, then selects a list of items, and observes stochastic outcomes of the selected items
	sequentially following the list order
	until some problem-specific stopping criterion happens.
In online recommendation, the stopping criterion might be the first item a user clicks; in networking, the stopping criterion might be the first edge blocked in a path. 
We further consider position discounts in the list order, which means the agent's reward may
	be discounted depending on the position where the stopping criterion is met.
We formulate this problem in a general contextual combinatorial cascading bandit setting.
We then propose a UCB-type algorithm to solve our problem, C$^3$-UCB, and prove an $n$-step regret bound $\tilde{O}(\sqrt{n})$ in both the general setting and two special cases. 
Our work generalizes existing studies in several directions, including contextual information,
	position discounts, and a more general cascading bandit model.
Our experiments on synthetic and real datasets demonstrate the advantage of involving contextual information and position discounts.

\end{abstract}

\section{Introduction}

Multi-armed bandit (MAB) has been extensively studied in statistics and machine learning. 
The problem can be formulated as a system of $K$ base arms whose rewards are random samples from unknown distributions with unknown means.
 The learning agent pulls one arm every time and tries to minimize the cumulative regret, 
 the difference in cumulative rewards between always pulling the best arm in expectation and
 playing according to the agent's strategy.
The problem of MAB has to deal with the trade-off between exploitation (pulling the best empirical arm) and exploration (trying other arms which are not sufficiently pulled).

Recently, stochastic combinatorial bandit started to draw many attention \cite{gai2012combinatorial,chen2013combinatorial,lin2014combinatorial,GMM14,chen2014pure,KWAEE14,kveton2014tight,kveton2015cascading,kveton2015combinatorial,lin2015online,Combes2015}. 
At every time step, a learning agent chooses a subset of ground items (super arm) under certain combinatorial constraints. There are several different kinds of feedback: (1) bandit feedback, where the learning agent can only obtain the reward of chosen super arm; (2) semi-bandit feedback, where the learning agent can also obtain the stochastic outcomes of the all base arms constituting the chosen super arm; 
(3) cascading feedback, where the learning agent can obtain the reward of the chosen super arm and the weights of some base arms in the chosen super arm, according to some problem-specific stopping criterion for observation.

Cascading feedback model fits into many real application scenarios.
For example, in online or mobile recommendation, it is a typical practice that an ordered list
	(instead of a single item) is recommended to a user, who
	usually goes over the list based on the recommended order and selects one of her interest to
	click through.
Thus it is reasonable to assume that items before the clicked item is of no interest to the user
	while user's interest to items after the clicked item is unclear.
Another example is in networking routing, where the agent needs to select a routing path that is least
	likely of being blocked.
To determine whether a path is blocked, the agent checks from the source node until meeting a blocked edge. 
The edges before the blocked one are unblocked and the edges after the blocked one are unobserved.
In both of the above applications, we observe the feedbacks of a prefix of items in the chosen ordered list. The bandit problem with cascading feedback has been studied in \cite{kveton2015cascading,kveton2015combinatorial}. 
In this paper, we generalize their work in several directions to cover more realistic scenarios.

First, we incorporate contextual information into cascading bandits.
In online recommendation, contextual information includes various user and item information, 
	and user behaviors in different contexts are different.
Therefore utilizing contextual information is crucial for personalized recommendation.
In network routing, temporal and spacial contextual information may also be useful to
	determine better routing paths.
Therefore, we incorporate contextual information into the cascading bandit formulation.
Second, cascading bandits studied in \cite{kveton2015cascading,kveton2015combinatorial}
	treats all positions in the cascading list equally, but in applications different positions
	may bring different rewards.
For example, in online recommendation, we usually prefer users to find their interested item
	in the list as early as possible to increase user satisfaction.
In network routing, we may prefer to hit a blocked edge as late as possible.
To model these preference, we introduce position discounts to the cascading bandit, and
	we show through experiments that incorporating position discounts may significantly
	improve the learning result.
Finally, we generalize the reward functions of \cite{kveton2015cascading,kveton2015combinatorial}
	that are based on disjunctive and conjunctive binary random variables to more general
	non-linear reward functions satisfying monotonicity and Lipschitz continuity conditions.
This generalization allows us to covers new realistic scenarios not covered in
	\cite{kveton2015cascading,kveton2015combinatorial}, and we give such an example
	in Section~\ref{sec:diss}.

%Personalized recommendation is a key feature of online recommendations to improve user satisfaction by providing attractive contents to meet specific user's need. Personalization includes storing user features and historical information, managing content set, analysis of user behaviors and providing suitable content to the user. Usually, both users and contents are represented by feature vectors. User features may contain basic attributes, like age, gender and nationality, and historical behaviors. Content features may contain categories and descriptive information. Because the views of different users on the same content can vary significantly, exploration and exploitation have to be exercised at an individual level and be able to cross different contents. When each item is represented by a feature vector that can be observed by the learning agent, the problem is known as contextual combinatorial bandit, which is developed in recommender systems and studied in \cite{qin2014contextual}.

%In this paper, we propose a general framework for contextual cascading bandits encompassing the applications of recommendations and network routing problems as special cases. In addition, we take position discounts into consideration. In recommendations, different positions usually give different values. In advertisement display, there are different positions in a website, the front middle, the left/right column and the bottom, and the positions in the front middle might charge highest price. Users are most satisfied if the first returned result is what they want. We introduce the position discounts to model this variant of rewards.

%Putting context, position discounts, and generalized reward function together, we call our learning model the {\it contextual combinatorial cascading bandits}. We formally define this model, provide a complete algorithm and its regret analysis, and conduct experiments on both synthetic and real-world recommendation and network routing datasets to validate our approach.

%The main contributions of this paper can be summarized as follows: (1) we propose a general framework  that incorporates contextual information, position discounts, and general reward function to cover many real-world application scenarios; (2) we propose $C^3$-UCB, an efficient algorithm for contextual combinatorial cascading bandit, with regret bound as $\tilde{O}(\sqrt{n})$ for $n$ rounds; (3) we evaluate this algorithm on synthetic and real-world datasets and the results show that our algorithm effectively incorporates contextual information and position discounts with significant performance gains.

The organization of this paper is as follows: section 2 compares our work with previous works; section 3 presents the formulation of {\it contextual combinatorial cascading bandits} problem with general reward function and position discounts; section 4 gives our algorithm and main results both on general reward fucntions and two special reward functions; section 5 demonstrates the experimental results; and section 6 gives a final conclusion.

%\wei{The last two paragraphs have some repetition. Is there any technical part in the algorithm and analysis that we would like to mention in the introduction? So far the intro is only about conceptual additions to the model, and nothing about technical challenges and our solution. It is ok if conceptual contribution is what we want to emphasize, but if there is any technical contribution we can say, we can also say it here.}

%\shuai{Our proofs build on previous work but also generalize their proof in our setting. Two challenges have been dealt with, partial observability on contextual bandits and general reward function with position discounts.  }

\begin{table}
	\renewcommand{\arraystretch}{1.1}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
		\hline
		 &context &cascading &position  & general \\
		 &         &          &discount & reward \\
		\hline
		CUCB &no &yes &no & yes \\
		\hline
		C$^2$UCB &yes &no &no & yes  \\
		\hline
		CombCascade	&no &yes &no  & no \\
		\hline
		C$^3$-UCB(ours) &yes &yes &yes & yes \\
		\hline
	\end{tabular}
	\caption{Comparisons of our setting with previous ones }
	\label{table:outline}
\end{table}

\section{Related Work}

We first discuss several studies most relevant to our work.
Table \ref{table:outline} summarizes the different settings of these work, 
	which we will explain in details next.
A comparison of regret bounds by different methods is deferred to Section \ref{sec:diss},
	after we provide the full result of our regret bound.

The work of \cite{kveton2015cascading} introduced the cascading model to the multi-armed bandit framework with disjunctive objective, where the set of feasible actions form a uniform matroid, and the reward of an action is $1$ if there is at least one ``good'' item in the list. 
The combinatorial cascading bandits of \cite{kveton2015combinatorial} (refered to as 'CombCascade' in Table \ref{table:outline}) generalized the framework, where each feasible action is a subset of ground items under combinatorial constraints; it studied a problem of combinatorial bandits with conjunctive objective, where each feasible action is a chain of items, and reward is $1$ if all items are "good". 
Our work generalizes these model and involves both contextual information and position discounts.

The experiments of \cite{kveton2015cascading} found that recommending items in increasing order of their UCBs (meaning the items with lower preference comes first) has a better performance than in decreasing order of their UCBs.
This unnatural result is perhaps because the order of items does not affect the reward and thus
	putting low preference items first and high preference items last may result in more feedbacks
	in the cascade.
The position discounts introduced in this paper forces the recommended set in the decreasing order 
	of UCBs, which is more realistic.
The work of \cite{combes2015learning} considered a similar cascading model to that of \cite{kveton2015cascading} with a particular case of contextual information, where at each time step, the category, which the user belongs to, is known and the user is recommended with $K$ items from a user-related group of items. Under this setting, they considered position discounts.
In our setting, users and items are represented by general feature vectors.

Our work considers contextual combinatorial bandits with cascading feedback and nonlinear reward. The paper \cite{li2010contextual} studied multi-armed bandit with context information, which recommends one item a time. A recent work \cite{qin2014contextual} (refered to as 'C$^2$UCB' in Table \ref{table:outline}) introduced contextual combinatorial bandit with semi-bandit feedback and nonlinear reward. Their feedback is more informative than ours. We relax the feedback from semi-bandit to cascading, yet our algorithm achieves the same order of regret bound.

There have been several pieces of work on partial monitoring MAB, but the results are not applicable to our setting. The papers of \cite{agrawal1989asymptotically,bartok2012adaptive} studied partial monitoring problems but their algorithms are inefficient in our setting. The work \cite{lin2014combinatorial} studied combinatorial partial monitoring problem and their feedback is a fixed (with respect to chosen action) combination of weights. In our setting, if chosen action is fixed, the number of observed items varies with the user's behavior, thus the combination constituting feedback is not fixed.

\citet{chen2015combinatorial} consider general reward functions and probabilistic triggering of base arms
	in the  stochastic combinatorial semi-bandit model (refered to as 'CUCB' in Table \ref{table:outline}).
Cascading bandits can be viewed as a particular type of probabilistic triggering, where
	only the head of the list is deterministically triggered and all the remaining items are
	probabilistically triggered.
However, their work does not consider contextual information, and it is also difficult to
	incorporate position discounts in a general triggering model.

% and / or case	
% \section{Problem Formulation}

% We formulate the problem of {\em contextual combinatorial cascading bandit} as follows. 
% Suppose we have a finite set  $E=\{1,...,L\}$ of $L$ \textit{ground items},  also referred to as {\em base arms}. 
% Let $\Pi^k=\{(a_1,...,a_k): a_1,...,a_k \in E, a_i \neq a_j \text{ for any } i \neq j\}$ be the set of all $k$-tuples of distinct items from $E$. 
% We call each of such tuples an {\em action} of length $k$. We will use $|A|$ to denote the length of an action $A$.
% Let $\Pi^{\leq K}= \cup_{k=1}^K \Pi^{k}$ denote the set of all actions with length at most $K$, and let $\cS \subseteq \Pi^{\leq K}$ be the set of \textit{feasible actions} with length at most $K$.

% As a convention, we always use boldface symbols to represent random variables.

% At time $t$, feature vectors $x_{t,a} \in \RR^{d \times 1}$ with $\norm{x_{t,a}}_2 \leq 1$ for every base arm $a \in E$ are revealed to the learning agent; the feature vectors combine both the information of the user and the corresponding base arm. \shengyu{Add a standard/typical example (just one sentence) to illustrate the combination?} Then the learning agent recommends a feasible action $\bA_t=(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t) \in \cS$ to a user, where notice that the length of the action is not fixed. Starting from the first item $\ba_{1}^t$, the user checks the items $(\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t)$ one by one in that order. The checking process stops if the user clicks \shengyu{selects?} on one item or has checked all items without clicking \shengyu{selecting?} anyone. We use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ as the {\em weight} of base arm $a$ at time $t$ to indicate whether the user has clicked on $a$ or not.

% Let $\cH_t$ denote the history before the learning agent chooses action at time $t$.
% Thus $\cH_t$ contains all information and observations at time $s < t$ and context information at time $t$. Given history $\cH_t$, we assume $\bw_{t,a}$'s are mutually independent and satisfy
% \begin{equation} % eq:expectation
%   \label{eq:expectation}
%   \EE[\bw_{t,a} | \cH_t] = \theta_{\ast}^{\top} x_{t,a},
% \end{equation}
% where $\theta_{\ast}$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_{\ast}}_2 \leq 1$ and $0 \leq \theta_{\ast}^{\top} x_{t,a} \leq 1$ for all $t, a$. 
% We define random variable $\bO_t$ as the number of observed base arms in $\bA_t$, that is, for some $k=1,2,\ldots, \abs{\bA_t}$, 
% $$
%   \bO_t = k, \text{ if } \bw_t(\ba_j^t)=0, \forall\, j < k \text{ and } \bw_t(\ba_k^t) = 1,
% $$
% or 
% $$
%   \bO_t = \abs{\bA_t}, \text{ if }\bw_t(\ba_j^t) = 0, ~~ \forall\, j \leq \abs{\bA_t}.
% $$
% At the end of every time $t$, the agent observes the first $\bO_t$ items of $\bA_t$. We say that item $a$ is {\it observed} if $a = \ba_k^t$ for some $k \leq \bO_t$. 
% Thus, $\cH_t$ consists of $\{x_s, \bA_s = (\ba_{1}^s,...,\ba_{\abs{\bA_s}}^s), \bO_s, \bw_s(\ba_k^s),x_t \mid  k \in[\bO_s], 1 \le s<t \}$.

% The agent receives some reward if the user clicks on some item. Suppose we consider the position discount: if the user clicks on the $k$-th item, then the learning agent receives reward $0 \leq \gamma_k \leq 1$. Usually the importance of the positions is decreasing: the first position is most important. So it is a reasonable assumption that
% $$
%   1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_K \geq 0.
% $$
% At the end of time $t$, the learning agent observes $\bO_t, \bw_t(\ba_k^t), k \leq \bO_t$ and receives reward
% $$
%   \br_t = \bw_t(\ba_{\bO_t}^t) \gamma_{\bO_t} = \bigvee_{k=1}^{\abs{\bA_t}} \gamma_k \bw_t(\ba_k^t).
% $$
% where we use the notation that $\bigvee_{k=1}^n a_k = \max_{1 \leq k \leq n} a_k$. Notice that the order of $\bA_t$ affects both the feed back and the reward.

% Now let us introduce a function $f$ on $A = (a_1,...,a_{\abs{A}}) \in \cS, w = (w(1),...,w(L))$ by
% \begin{align}
%   &f : \cS \times [0,1]^E \to [0,1] \nonumber \\
%   &f(A,w) = \sum_{k = 1}^{\abs{A}} \gamma_{k} \prod_{i=1}^{k-1} (1 - w(a_i)) w(a_k).
%   \label{eq:functionf}
% \end{align}
% Then we have $\br_t = f(\bA_t, \bw_t)$ and $\EE[\br_t] = f(\bA_t, \theta_{\ast}^{\top}x_t)$ where $x_t = (x_{t,1}, \cdots, x_{t,L}) \in \RR^{d \times L}$. Let 
% \begin{align*}
%   A_t^{\ast} &= \argmax_{A\in \cS} f(A,\theta_{\ast}^{\top}x_t),\\
%   f_t^{\ast} &= f(A_t^{\ast}, \theta_{\ast}^{\top}x_t).
% \end{align*}
% The goal of our learning algorithm is to minimize the expected cumulative regret in $T$ steps
% $$
%   R(T) = \EE\left[\sum_{t=1}^T R(t, \bA_t)\right].
% $$
% where $R(t, A) = f_t^{\ast} - f(A, \theta_{\ast}^{\top}x_t)$.

% The above formulation is on the disjunctive objective, that is, at a time step the agent stops as soon as she reveals a base arm at a position $k$ in the sequence with weight $1$ and she receives reward $\gamma_k$ as the result. 
% Similarly, we could also consider the case of conjunctive objective. 
% We also use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ to indicate the weight of item $a$ at time $t$ satisfying Equation \eqref{eq:expectation}). 
% The learning agent observes all items until the first one with weight $0$. 
% Suppose we also consider partial reward: if the $k$-th item is the first item with weight $0$, then the learning agent receives reward $\gamma_k'$; if all items have weight $1$, the learning agent receives reward $1$. 
% The more items the learning agent reveals, the more reward the agent should receive. It is reasonable to assume that
% $$
%   0 = \gamma'_1 \leq \gamma'_2 \leq \cdots \leq \gamma'_K \leq 1.
% $$
% At the end of time $t$, the learning agent observes $\bO_{\wedge, t}, \bw_t(\ba_k^t), k \leq \bO_{\wedge, t}$ and receives reward $\br_{\wedge, t}$:
% \begin{align*}
%   \br_{\wedge, t} &= \begin{cases}
%     \gamma_{\bO_{\wedge, t}}'  &\text{if } \bw_t(\ba_{\bO_{\wedge, t}}^t) = 0,\\
%     1 &\text{if } \bw_t(\ba_{\bO_{\wedge, t}}^t) = 1,
%   \end{cases}\\
%   &=\begin{cases}
%     \gamma_{k}'  &\text{if } \bw_t(\ba_{i}^t) = 1, i < k, \bw_t(\ba_{k}^t) = 0,\\
%     1 &\text{if } \bw_t(\ba_{i}^t) = 1, i\leq \abs{\bA_t}.
%   \end{cases}
% \end{align*}

% If we define a function $f_{\wedge}$ on $A = (a_1, \ldots, a_{\abs{A}}) \in \cS, w = (w(1), \ldots, w(L))$ by
% \begin{align}
%   &f_{\wedge} : \cS \times [0,1]^E \to [0,1] \nonumber \\
%   &f_{\wedge}(A,w) = \sum_{k = 1}^{\abs{A}} \gamma_k' \prod_{i = 1}^{k - 1} w(a_i)(1 - w(a_k)) + \prod_{i=1}^{\abs{A}}w(a_i),
%   \label{eq:functionfstar}
% \end{align}
% then we have $\br_{\wedge, t} = f_{\wedge}(\bA_t, \bw_t)$ and $\EE[\br_{\wedge, t}] = f_{\wedge}(\bA_t, \theta_{\ast}^{\top}x_t)$ where $x_t = (x_{t,1}, \cdots, x_{t,L}) \in \RR^{d \times L}$. To simplify notations, we assume 
% $$
%   \gamma_k' = 1 - \gamma_k.
% $$
% Then it holds that
% \begin{equation} %eq:ConDisRelation
%   \label{eq:ConDisRelation}
%   f_{\wedge}(A, w) = 1 - f(A, 1 - w).
% \end{equation}
% Let 
% \begin{align*}
%   A_{\wedge, t}^{\ast} &= \argmax_{A\in \cS} f(A,\theta_{\ast}^{\top}x_t),\\
%   f_{\wedge, t}^{\ast} &= f(A_t^{\ast}, \theta_{\ast}^{\top}x_t).
% \end{align*}
% For the conjunctive objective, the goal of our learning algorithm is to minimize the expected cumulative regret in $T$ steps
% $$
%   R_{\wedge}(T) = \EE\left[\sum_{t=1}^T R_{\wedge}(t, \bA_t)\right].
% $$
% where $R_{\wedge}(t, A) = f_{\wedge, t}^{\ast} - f_{\wedge}(A, \theta_{\ast}^{\top}x_t)$.

\section{Problem Formulation}

We formulate the problem of {\em contextual combinatorial cascading bandits} as follows. Suppose we have a finite set $E=\{1,...,L\}$ of $L$ \textit{ground items}, also referred to as {\em base arms}. Let $\Pi^k=\{(a_1,...,a_k): a_1,...,a_k \in E, a_i \neq a_j \text{ for any } i \neq j\}$ be the set of all $k$-tuples of distinct items from $E$; we call each of such tuples an {\em action} of length $k$. We will use $|A|$ to denote the length of an action $A$. Let $\Pi^{\leq K}= \cup_{k=1}^K \Pi^{k}$ denote the set of all actions with length at most $K$, and let $\cS \subseteq \Pi^{\leq K}$ be the set of \textit{feasible actions} with length at most $K$.

As a convention, we always use boldface symbols to represent random variables.

At time $t$, feature vectors $x_{t,a} \in \RR^{d \times 1}$ with $\norm{x_{t,a}}_2 \leq 1$ for every base arm $a \in E$ are revealed to the learning agent. Each feature vector combines the information of the user and the corresponding base arm. For example, suppose the user at time $t$ is characterized by a feature vector $u_t$ and the base arm $a$ has a feature vector $v_a$, then we can use $x_{t,a} = (u_t, v_a)$ , the concatenation of $u_t$ and $v_a$, as the combined feature vector of base arm $a$ at time $t$. Denote $x_t = (x_{t,a})_{a \in E}$ as all context information at time $t$. Then the learning agent recommends a feasible action $\bA_t=\{ \ba_{1}^t,...,\ba_{\abs{\bA_t}}^t \} \in \cS$ to the user. In the cascading feedback models, the user checks from the first item of recommended action and stops at the $\bO_t$-th item under some stopping criterion. For example in recommender systems, the stopping criterion might be that the user finds an attractive item. Then the learning agent observes the weights of first $\bO_t$ base arms in $\bA_t$, denoted by $\bw_t(\ba_k^t), k \leq \bO_t$.

Let $\cH_t$ denote the history before the learning agent chooses action at time $t$. Thus $\cH_t$ contains feedback information at all time $s < t$, as well as context information at time $t$. Each arm $a$ has a weight $\bw_t(a)$, representing the "quality" to the user at time $t$. Given history $\cH_t$, we assume that $\bw_{t,a}$'s are mutually independent random variables with expectation
\begin{equation} % eq:expectation
\label{eq:expectation}
\EE[\bw_{t}(a) | \cH_t] = \theta_{\ast}^{\top} x_{t,a},
\end{equation}
where $\theta_{\ast}$ is an unknown $d$-dimensional vector with the assumption that $\norm{\theta_{\ast}}_2 \leq 1$ and $0 \leq \theta_{\ast}^{\top} x_{t,a} \leq 1$ for all $t, a$. Denote by 
$$
w_{t,a} = \theta_{\ast}^{\top} x_{t,a}
$$
the expected weight of base arm $a$ at time $t$. We assume that each $\bw_{t}(a)$ is a random variable with $R$-sub-Gaussian tail, which means that for all $\lambda \in \RR$, 
$$
\EE[\exp(\lambda(\bw_{t}(a) - \theta_{\ast}^{\top} x_{t,a})) | \cH_t] \leq \exp(\lambda^2 R^2 / 2).
$$

Recall that the random variable $\bO_t$ is the number of observed base arms in $\bA_t$ and at the end of every time $t$, the agent observes the first $\bO_t$ items of $\bA_t$. We say that item $a$ is {\it observed} if $a = \ba_k^t$ for some $k \leq \bO_t$. Thus, $\cH_t$ consists of $\{x_s, \bA_s = (\ba_{1}^s,...,\ba_{\abs{\bA_s}}^s), \bO_s, \bw_s(\ba_k^s) \}_{k\leq \bO_s, s < t}$ and $\{x_t\}$.

We introduce position discounts $\gamma_k \in [0,1], k\leq K$. For example in website recommendations by a search engine, if the user selects the first website, the learning agent will receive reward $1$; and if the user selects the $k$-th website, the learning agent will receive reward $\gamma_k$.

The reward function on round $t$ is an application dependent function. We assume the expected reward of action $A$ is a function $f(A,w)$ of {\it expected weights} $w = (w(a))_{a \in E} \in [0,1]^{E}$ (at time $t$, $w = (w_t(a))_{a \in E} = (\theta_{\ast}x_{t,a})_{a \in E}) $ and satisfies the following two assumptions.

\paragraph{Monotonicity}
The expected reward function $f(A, w)$ is a non-decreasing with respect to $w$: for any $w,w'\in [0,1]^E$, if $w(a) \leq w'(a)$, we have $f(A, w) \leq f(A, w')$.

\paragraph{Lipschitz continuity}
The expected reward function $f(A, w)$ is Lipschitz continuous with respect to $w$ together with position discount parameters $\gamma_k, k\leq K$. In particular, there is a constant $B>0$ such that, for any $w,w' \in [0,1]^E$, we have
$$
	|f(A, w) - f(A, w')| \leq B \sum_{k=1}^{|A|} \gamma_k |w(a_k) - w'(a_k)|,
$$
where $A = (a_1, \ldots, a_{|A|})$.

The assumption of Lipschitz continuity gives an estimate of changes in reward function. To obtain a good estimation of reward function, the position $k$ with large $\gamma_k$ needs a good estimation of $w(a_k)$. This means the positions with large $\gamma_k$ are more important to the reward function. For example in website recommendations, the learning agent will receive highest reward if the user selects the first item, so the first position is most important.

Notice that we do not require the knowledge how $f(A, w)$ is defined. We only assume that the agent has access to an oracle $\cO_{\cS}(w)$ that outputs recommended action $A$. More precisely, an oracle $\cO_{\cS}(w)$ is called an $\alpha$-approximation oracle for some $\alpha \leq 1$, if on given input $w$, the oracle returns an action $A = \cO_{\cS}(w) \in \cS$ satisfying $f(A,w) \geq \alpha f(A^*,w)$ where $A^* = \argmax_{A\in\cS} f(A, w)$.

The $\alpha$-regret of action $A$ on time $t$ is
$$
R^{\alpha}(t, A) = \alpha f_t^{\ast} - f(A, w_t),
$$
where $f_t^{\ast} = f(A_t^*, w_t)$, $A_t^* = \argmax_{A\in\cS} f(A, w_t)$ and $w_t = (\theta_{\ast}^{\top}x_{t,a})_{a \in E}$. Our goal is to minimize the $\alpha$-regret
$$
R^{\alpha}(n) = \EE[ \sum_{t=1}^n R^{\alpha}(t, \bA_t)].
$$

\section{Algorithms and Results}

\subsection{Algorithm}

\begin{algorithm} %C^3-UCB algorithm
	\caption{C$^3$-UCB}
	\label{alg:ConComCascade}
	\begin{algorithmic}[1]
		\STATE {Parameters:}
		\STATE {\qquad $\{\gamma_k \in [0,1]\}_{k\leq K}; \delta = \frac{1}{\sqrt{n}}; \lambda \geq C_\gamma = \sum_{k=1}^{K} \gamma_k^2$}
		\STATE {Initialization:}
		\STATE {\qquad $\hat{\theta}_0 = 0, \bbeta_0(\delta) = 1, \bV_0 = \lambda I, \bX_0=\emptyset, \bY_0=\emptyset$}

		\FORALL {$t=1,2,\ldots$}
		\STATE {Obtain context $x_{t,a}$ for all $a\in E$}
		\STATE {$\forall a\in E$, compute}
		\STATE {~~ $\bU_t(a) = \min\{\hat{\theta}_{t-1}^{\top}x_{t,a} + \bbeta_{t-1}(\delta)\norm{x_{t,a}}_{\bV_{t-1}^{-1}}, 1\}$}

		\STATE {//Choose action $\bA_t$ using UCBs $\bU_t$}
		\STATE { $\bA_t=(\ba_1^t,...,\ba_{\abs{\bA_t}}^t) \leftarrow \cO_{\cS}(\bU_t)$ } 
		\STATE {Play $\bA_t$ and observe $\bO_t, \bw_t(\ba_k^t), k\in[\bO_t]$}

		\STATE {//Update statistics}
		\STATE {$\bV_{t} \leftarrow \bV_{t-1} + \sum_{k=1}^{\bO_t} \gamma_k^2 x_{t, \ba_k^t}x_{t, \ba_k^t}^{\top}$}
		\STATE {$\bX_t \leftarrow [\bX_{t-1}; ~\gamma_1 x_{t, \ba_1^t}^{\top};  ~\ldots; ~\gamma_{\bO_t} x_{t, \ba_{\bO_t}^t}^{\top}]$}
		\STATE {$\bY_t \leftarrow [\bY_{t-1}; ~\gamma_1 \bw(\ba_1^t); ~\ldots; ~\gamma_{\bO_t} \bw_t(\ba_{\bO_t}^t)]$}
		\STATE {$\hat{\theta}_t \leftarrow (\bX_t^\top \bX_t + \lambda I)^{-1} \bX_{t}^\top \bY_t$}
		\STATE {$\bbeta_{t}(\delta) \leftarrow R \sqrt{\ln(\det(\bV_{t})/(\lambda^d \delta^2))} + \sqrt{\lambda}$ }
		\ENDFOR {~~$t$}
	\end{algorithmic}
\end{algorithm}

%At time $t$, we denote $\EE_t[\cdot] = \EE[\cdot | \cH_t]$, where $\cH_t$ is the history a time $t$ before the learning agent chooses her action.

By Equation \eqref{eq:expectation}, we have 
$$
  \EE[(\gamma_k \bw_{s,\ba_k^s}) | \cH_{s}] = \theta_*^{\top} (\gamma_k x_{s,\ba_k^s}).
$$
Using the ridge regression of data 
$$
  \{(\gamma_k x_{s,\ba_k^s}, \gamma_k \bw_{s,\ba_k^s})\}_{k \in[\bO_s], s\in[t]},
$$
we get an $l^2$-regularized least-squares estimate of $\theta_*$ with regularization parameter $\lambda > 0$:
\begin{equation}
  \hat{\theta}_t = (\bX_t^{\top}\bX_t + \lambda I)^{-1} \bX_t^{\top} \bY_t,
\end{equation}
where $\bX_t \in \RR^{(\sum_{s=1}^{t}\bO_s) \times d}$ is the matrix whose rows are $\gamma_k x_{s,\ba_k^s}^{\top}$ and $\bY_t$ is the column vector whose elements are $\gamma_k \bw_s(\ba_k^s)$, $k \in[\bO_s], s\in[t]$. Let
$$
  \bV_t = \bX_t^{\top} \bX_t + \lambda I = \lambda I + \sum_{s=1}^{t} \sum_{k=1}^{\bO_s} \gamma_k^2 x_{s,\ba_k^s}x_{s,\ba_k^s}^{\top}.
$$
Then $\bV_t \in \RR^{d \times d}$ is a symmetric positive definite matrix. For any symmetric positive definite matrix $V \in \RR^{d \times d} $ and any vector $x \in \RR^{d \times 1}$, we define the $2$-norm of $x$ based on $V$ to be $\norm{x}_V = (x^{\top} V x)^{\frac{1}{2}}$. Next we build on a good estimate of the difference between $\hat{\theta}_t$ and $\theta_*$ by Theorem 2 in \cite{abbasi2011improved}, which states the following results.

\begin{lemma} $\mathrm{\cite{abbasi2011improved}}$ %estimate of theta
  \label{thm:theta_estimate}
  Let 
  \begin{equation}
    \bbeta_{t}(\delta) = R\sqrt{\ln\left(\frac{\det(\bV_{t})}{\lambda^d \delta^2}\right)} + \sqrt{\lambda}. \label{eq:definebeta}
  \end{equation}
  Then for any $\delta > 0$, with probability at least $1 - \delta$, for all $t > 0$, we have
  \begin{equation}
    \label{eq:estimateTheta}
    \norm{\hat{\theta}_t - \theta_{\ast}}_{\bV_{t}} \leq \bbeta_{t}(\delta).
  \end{equation}
\end{lemma}

This theorem states that with high probability, the estimate $\hat{\theta}$ lies in the ellipsoid centered at $\theta_*$ with confidence radius $\bbeta_t(\delta)$ under $\bV_t$ norm. 
Building on this, we can define an upper confidence bound of the expected weight of base arm $a$ by
\begin{equation}
  \label{eq:defU}
  \bU_t(a) = \min\left\{\hat{\theta}_{t-1}^{\top}x_{t,a} + \bbeta_{t-1}(\delta)\norm{x_{t,a}}_{\bV_{t-1}^{-1}}, 1 \right\}.
\end{equation}

The fact that $\bU_t(a)$ is an upper confidence bound of expected weight $w_{t, a} = \theta_*^{\top} x_{t,a}$ is proved in the following lemma.
\begin{lemma} %estimate of U
  \label{lem:estimateU}
  When Eq.\eqref{eq:estimateTheta} holds for time $t-1$, we have
  $$
    0 \leq \bU_t(a) - w_{t,a} \leq 2 \bbeta_{t-1}(\delta)\norm{x_{t,a}}_{\bV_{t-1}^{-1}}.
  $$
\end{lemma}
\begin{proof}
	$w_{t,a} = \theta_*^{\top} x_{t,a}$. By Holder's inequality,
	\begin{align*}
		\abs{\hat{\theta}_{t-1}^{\top}x_{t,a} - \theta_{\ast}^{\top}x_{t,a}} &\leq \norm{\hat{\theta}_{t-1} - \theta_{\ast}}_{\bV_{t-1}} \norm{x_{t,a}}_{\bV_{t-1}^{-1}} \\
		&\leq \bbeta_{t-1}(\delta)\norm{x_{t,a}}_{\bV_{t-1}^{-1}}.
	\end{align*}
	Then because
	\begin{align*}
		0 &\leq (\hat{\theta}_{t-1}^{\top}x_{t,a} + \bbeta_{t-1}(\delta)\norm{x_{t,a}}_{\bV_{t-1}^{-1}}) - \theta_{\ast}^{\top}x_{t,a} \\
		&\leq 2 \bbeta_{t-1}(\delta)\norm{x_{t,a}}_{\bV_{t-1}^{-1}}
	\end{align*}
	and the definition of $\bU_t$ in \ref{lem:estimateU}, we can obtain the result.
\end{proof}

Our proposed algorithm, C$^3$-UCB, is described in Algorithm \ref{alg:ConComCascade}. First, the learning agent computes the upper confidence bounds (UCBs) $\bU_t \in [0,1]^{E}$ for the expected weights of all base arms in $E$. Second, the agent uses the computed UCBs, $\bU_t$, to select an action $\bA_t = \{\ba_{1}^t,..., \ba_{\abs{\bA_t}}^t \}$. Third, the user checks from the first base arm in $\bA_t$ and stops after checking the $\bO_t$-th base arm and the agents observes $\bw_t(\ba_k^t), k\leq \bO_t$. Then, the learning agent updates $\bV_t, \bX_t, \bY_t$ in order to get a newer estimate $\hat{\theta}_t$ of $\theta_*$ and new confidence radius $\bbeta_t(\delta)$. In this part, we use the notation $[A; B]$ to denote the matrix obtained by stacking A and B vertically like $\begin{pmatrix} A\\ B\end{pmatrix}$.

\subsection{Results}

\subsubsection{General Results}

To state our main theorem, we need some definitions. Let $p_{t, A}$ be the {\em probability of full observation of $A$}, that is the probability of observing all base arms of $A = (a_1, \ldots, a_{\abs{A}})$ at time $t$. Let $p^* = \min_{1 \leq t \leq T} \min_{A \in \cS} ~ p_{t, A}$ to be the minimal probability that an action could have all base arms observed over all time. The following is the main theorem on the regret achieved by our C$^3$-UCB algorithm.

\begin{theorem} %main theorem
	\label{thm:main}
	Suppose the expected reward function $f(A, w)$ is a function of expected weights and satisfies the requirements of monotonicity and Lipschitz continuity with bound $B$. Assume there is an $\alpha$-approximation oracle $\cO_{\cS}$. Then the $\alpha$-regret of our algorithm, C$^3$-UCB, satisfies
	\begin{align}
		R^{\alpha}(n) &\le \frac{\sqrt{2}B}{p^*} \sqrt{nKd\ln(1 + C_\gamma n/(\lambda d))}  \nonumber \\
		&\qquad \cdot \left(R\sqrt{\ln \left[(1 + C_\gamma n/(\lambda d))^d n \right] } + \sqrt{\lambda}\right) \nonumber \\
		&=O\left(\frac{dBR}{p^*} \sqrt{nK} \ln(C_\gamma n) \right),
	\end{align}
	where $R$ is the sub-Gaussian constant and $C_{\gamma} = \sum_{k=1}^K \gamma_k^2 \leq K$.
\end{theorem}
\begin{proof} [of Theorem \ref{thm:main}] %proof outline of main theorem
	The complete proof is in the Appendix. The main idea is to reduce the regret at time $t$, $R^{\alpha}(t, \bA_t)$, to the case that all items in $\bA_t$ have been observed. This reduction is challenging for two reasons. First, our setting has contextual information. Second, we consider position discounts $\gamma_k, k \leq K$.

	Assume $\bA_t = (\ba_1^t, \ldots, \ba_{|\bA_t|}^t)$. Then by monotonicity and Lipschitz continuity, we have
	$$
		R^{\alpha}(t, \bA_t) \leq B \sum_{k=1}^{|\bA_t|} \bbeta_{t-1}(\delta) \gamma_k \| x_{t, \ba_k^t} \|_{\bV_{t-1}}.
	$$
	However, it is hard to estimate an upper bound for $\sum_{t=1}^T \sum_{k=1}^{\abs{\bA_t}} \norm{ x_{t, \ba_k^t} }_{ \bV_{t-1}^{-1} }$ because $\bV_t$ only contains information of observed base arms. Thus we need the following deduction.
	\begin{align*}
    	&\EE[R^{\alpha}(t, \bA_t)|\cH_t] \\
    	&\qquad = \EE \left[ \left. R^{\alpha}(t, \bA_t) \EE \left[ \left. \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t = \abs{\bA_t}\} \right| \bA_t \right]  \right| \cH_t \right]\\
    	&\qquad \leq \frac{1}{p^*} \EE \left[ \left. R^{\alpha}(t, \bA_t) \bOne\{\bO_t = \abs{\bA_t}\} \right| \cH_t \right].
    % &\EE_t[R_{\wedge}(t, \bA_t)] \\
    % &\qquad \EE \left[ \left. R_{\wedge}(t, \bA_t) \EE \left[ \left. \frac{1}{p_{\wedge, t, \bA_t}} \bOne\{\bO_t\geq\abs{\bA_t}\} \right| \bA_t \right]  \right| \cH_t \right] \\
    \end{align*}
    Then the cumulative regret is bounded by the discounted sum of the contextual information of observed base arms in $\bV$-norms. The next Lemma solves this issue.
    \begin{lemma}
    	\label{lem:XNormSumEst}
    	If $\lambda \geq C_{\gamma}$, then
    	$$
    	\sum_{s=1}^t \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \leq 2d\ln \left(1 + C_{\gamma}t/(\lambda d)\right),
    	$$
    \end{lemma}
    Then we can prove this theorem.
\end{proof}

\subsubsection{Disjunctive Objective}

In the problem of cascading recommendation, when recommended with a list of items $\bA_t = (\ba_{1}^t,...,\ba_{\abs{\bA_t}}^t)$, the user checks from the first item $\ba_{1}^t$ and then one by one in order. The checking process stops if the user selects one item or has checked all items without selecting anyone. In this case, the weight of base arm $a$ at time $t$, $\bw_{t,a} \in \{ 0, 1 \}$, is binary to indicate whether the user has clicked on $a$ or not. Then the random variable $\bO_t$ satisfies
$$
\bO_t = k, \text{ if } \bw_t(\ba_j^t)=0, \forall\, j < k \text{ and } \bw_t(\ba_k^t) = 1,
$$
or 
$$
\bO_t = \abs{\bA_t}, \text{ if }\bw_t(\ba_j^t) = 0, ~~ \forall\, j \leq \abs{\bA_t}.
$$

The agent receives some reward if the user clicks on some item: if the user clicks on the $k$-th item, then the learning agent receives reward $0 \leq \gamma_k \leq 1$. Usually the importance of the positions is decreasing: the first position is most important. So it is reasonable to assume that
$$
1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_K \geq 0.
$$
At the end of time $t$, the learning agent observes $\bO_t, \bw_t(\ba_k^t), k \leq \bO_t$ and receives reward
$$
\br_t = \bw_t(\ba_{\bO_t}^t) \gamma_{\bO_t} = \bigvee_{k=1}^{\abs{\bA_t}} \gamma_k \bw_t(\ba_k^t).
$$
where we use the notation that $\bigvee_{k=1}^n a_k = \max_{1 \leq k \leq n} a_k$. Notice that the order of $\bA_t$ affects both the feed back and the reward.

Now let us introduce a function $f$ on $A = (a_1,...,a_{\abs{A}}) \in \cS, w = (w(1),...,w(L))$ by
\begin{align}
&f : \cS \times [0,1]^E \to [0,1] \nonumber \\
&f(A,w) = \sum_{k = 1}^{\abs{A}} \gamma_{k} \prod_{i=1}^{k-1} (1 - w(a_i)) w(a_k). \label{eq:OrRewardFunc}
\end{align}
Then we have $\br_t = f(\bA_t, \bw_t)$ and $\EE[\br_t] = f(\bA_t, \theta_{\ast}^{\top}x_t) = f(\bA_t, w_t)$. So $f$ is the expected reward function and also a function of expected weights. It can be proved that $f$ satisfies the properties of monotonicity and Lipschitz continuity with bound $B=1$. (We leave all verifications in the appendix.) Notice that in this setting, given weights $w$, there is an oracle $\cO_{\cS}(w)$ that outputs the best action, which is the $K$ items with largest $w_i$ listed in decreasing order. So $\alpha = 1$ and we write $R(n)$ instead of $R^{\alpha}(n)$. Then we have the following corollary.
\begin{corollary}
	\label{cor:or}
	In the problem of cascading recommendation, the expected reward function is defined in Eq.~\eqref{eq:OrRewardFunc}, where $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_K \geq 0$. Then the $\alpha$-regret of our algorithm, C$^3$-UCB, satisfies
	\begin{align}
		R(n) &\le \frac{\sqrt{2}}{p^*} \sqrt{nKd\ln(1 + C_\gamma n/(\lambda d))}  \nonumber \\
		&\qquad \cdot \left(\sqrt{d\ln[(1 + C_\gamma n/(\lambda d))^d n]} + \sqrt{\lambda}\right) \nonumber \\
		&=O\left(\frac{d}{p^*} \sqrt{nK} \ln (C_\gamma n) \right).
	\end{align}
\end{corollary}

\subsubsection{Conjunctive Objective}
\label{sec:conj}

The above formulation is on the disjunctive objective, that is, at a time step the user stops as soon as she reveals a base arm at a position $k$ in the sequence with weight $1$ and the learning agent receives reward $\gamma_k$ as the result. Similarly, we could also consider the case of conjunctive objective. In this case, we can derive a slightly better result.

We also use Bernoulli random variable $\bw_{t}(a) \in \{0,1\}$ to indicate the weight of item $a$ at time $t$ satisfying Equation \eqref{eq:expectation}. The learning agent observes all items until the first one with weight $0$. Suppose we also consider partial rewards: if the $k$-th item is the first item with weight $0$, then the learning agent receives reward $1-\gamma_k$; if all items have weight $1$, the learning agent receives reward $1$.

The more items reveals, the more reward the agent should receive. It is reasonable to assume that
$$
1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_K \geq 0.
$$
At the end of time $t$, the learning agent observes $\bO_{t}, \bw_t(\ba_k^t), k \leq \bO_{t}$ and receives reward $\br_{t}$:
\begin{align*}
	\br_{t} &= \begin{cases}
		1 - \gamma_{\bO_{t}}  &\text{if } \bw_t(\ba_{\bO_{t}}^t) = 0,\\
		1 &\text{if } \bw_t(\ba_{\bO_{t}}^t) = 1,
	\end{cases}\\
	&=\begin{cases}
		1-\gamma_{k}  &\text{if } \bw_t(\ba_{i}^t) = 1, i < k, \bw_t(\ba_{k}^t) = 0,\\
		1 &\text{if } \bw_t(\ba_{i}^t) = 1, i\leq \abs{\bA_t}.
	\end{cases}
\end{align*}

If we define a function $f$ on $A = (a_1, \ldots, a_{\abs{A}}) \in \cS, w = (w(1), \ldots, w(L))$ by
\begin{align}
	&f: \cS \times [0,1]^E \to [0,1] \nonumber \\
	&f(A,w) = \sum_{k = 1}^{\abs{A}} (1 - \gamma_k) \prod_{i = 1}^{k - 1} w(a_i)(1 - w(a_k)) + \prod_{i=1}^{\abs{A}}w(a_i),
\label{eq:AndRewardFunc}
\end{align}
then we have $\br_{t} = f(\bA_t, \bw_t)$ and $\EE[\br_{t}] = f(\bA_t, \theta_{\ast}^{\top}x_t)$ where $x_t = (x_{t,1}, \cdots, x_{t,L}) \in \RR^{d \times L}$.

In the problem of network transportations or sequential task processing, $p^*$ defined above might be very small. So we have the following improvements. Suppose we also consider position discount here.  
\begin{theorem} %main theorem
	\label{thm:and}
	Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_{K} \geq 0$ and $\gamma_K \geq 1 - \frac{1}{4} f^{\ast}$, where $f^* = \min_{t} f_t^{\ast}$. Let regularization parameter $\lambda$ satisfy $\lambda \geq C_\gamma$, where $C_\gamma = \sum_{k=1}^{K} \gamma_k^2$. Then the $\alpha$-regret of our algorithm, C$^3$-UCB for the conjunctive objective, satisfies
	\begin{align}
		R(n) \le& \frac{\sqrt{128}}{f^{\ast}} \sqrt{nKd\ln(1 + C_\gamma n/(\lambda d))} \nonumber \\
		&\qquad \cdot \left(\sqrt{d\ln[(1 + C_\gamma n/(\lambda d)^d n] } + \sqrt{\lambda}\right) \nonumber \\
		=&O\left(\frac{d}{f^{\ast}}\sqrt{nK} \ln(C_\gamma n) \right).
	\end{align}
\end{theorem}
\begin{proof}
	We will prove it in the appendix. The main idea is that we prove the expected reward function satisfies the properties of monotonicity and Lipschitz continuity first. And then use the following lemma to replace $\bA_t$ with a prefix of $\bA_t$.
	\begin{lemma}
		\CLemmaPrefixExi
	\end{lemma}

	Therefore we can get
	\CEqDeltaEstAnd

	Then similar to the proof of Theorem \ref{thm:main} and by the Lemma \ref{lem:XNormSumEst}, we can prove this theorem.
\end{proof}

\subsection{Discussions}
\label{sec:diss}

In the above, we formulate the general setting of contextual combinatorial cascading bandit with position discount. This setting generalizes several existing results.

If our setting restricts to the setting of combinatorial cascading bandit \cite{kveton2015combinatorial} which coincides with our section \ref{sec:conj} without contextual information and with position discounts $\gamma_k = 1$, our regret bound is of order $O\left(\frac{L}{f^*}\sqrt{nK}\ln(n)\right)$ which has an additional term $\sqrt{\frac{L\ln(n)}{f^*}}$ than their result. The reason is that the best regret bound for stochastic MAB is $O(\frac{\ln(n)}{\Delta})$, which has been achieved by a lot of methods, while the best regret bound for linear bandits is $O\left(\frac{\ln^2(n)}{\Delta}\right)$ by \cite{abbasi2011improved} and we use the technique of linear bandits to derive a confidence ellipsoid of $\theta_{\ast}$.

If our setting restricts to contextual combinatorial bandits \cite{qin2014contextual}, where there are no triggering and no position discounts, so every base arm in action $\bA_t$ will be observed. The random variable $\bO_t$ of observed base arms should be $|\bA_t|$. The probability of full observation of $A$, $p_{t, A}$, should be $1$, so $p^{\ast} = 1$. And the position discounts $\gamma_k$ are all $1$. Then our regret bound is of the same order with theirs. Notice that if a reward function satisfies our assumptions of monotonicity and Lipschitz continuity, then the reward function satisfies their assumptions with the bound $C$ for Lipschitz continuity be $\sqrt{K}$.

Our algorithm can also apply to the setting of \cite{chen2013combinatorial} with context information. However, the probability $p^{\ast}$ in our results should be the probability that all base arms in $\tilde{S}$ (in their notations) will be triggered.

Our setting generalizes several existing works but does not limit to them. Consider in network routing problem, the observed information would be the latency on specific edges. The latency, or {\it weight}, on an edge is an exponential variable with mean $\theta_{\ast}^{\top}x$. Suppose we treat an edge as blocked if the latency on it is larger than some tolerance $\tau$. And we wait the edge to respond for at most $2\tau$ time. Then the real observed latency is new random variable from a cut-off exponential distribution with mean $\theta_{\ast}^{\top}x$. Then the edge is unblocked, or with reward $1$, if the new random vairable is less than the tolerance $\tau$ and the expectation is a function, say $g$, of $\theta_{\ast}^{\top}x$. It is not difficult to check that $g$ satisfies the properties of monotonicity and Lipschitz continuity. And the expected reward of a path is a composition of the expected reward function of conjunctive objective, defined in Eq.~\eqref{eq:AndRewardFunc}, and the function $g$. (In original formulation, $g$ is the identity function, where $g(w) = w$.) Such an expected reward function satisfies the properties of monotonicity and Lipschitz continuity and can not be represented in conjunctive/disjunctive objectives, like in \cite{kveton2015cascading,kveton2015combinatorial}.

%Our setting generalizes several existing works but not limited to them. If the expected reward $g$ of a base arm $a$ has the property of monotonicity and Lipschitz continuity with respect to $\theta_{\ast}^{\top}x_{a}$, then the conjunctive, or disjunctive, objective of those expected rewards can not be covered in

%in medical treatment, a patient will be recommended with a series of treatments. The patient has to take the treatments one by one, so the series of treatment actually is a list. Suppose the effect, after taking a treatment, is represented by a Beta random variable $\bw$ with a mean $w = \theta_*^{\top}x$. $\theta_*$ is unknown linear coefficients and $x$ denotes the context, which might include the conditions of the patient and the properties of the treatment. We say that the treatment is successful if the effect $\bw$ is larger than a threshold $\tau$. So a treatment $a$ is successful with probability
% \begin{equation}
% \label{eq:betaTau}
% 1 - \frac{B(\tau; \alpha, \beta_0)}{B(\alpha, \beta_0)},
% \end{equation}
% where $B$ is the incomplete Beta function, $\alpha = \frac{\beta_0 w}{1- w} > 1$ and $\beta_0 > 1$ is a fixed number. We say that a series of treatment $A = (a_1, \ldots, a_{|A|})$ is successful (value $1$) if and only if all treatments in $A$ are successful. So the expected reward function is
% $$
% f(A, w) = \prod_{k=1}^{|A|} (1 - \frac{B(\tau; \alpha_{a_k}, \beta_0)}{B(\alpha_{a_k}, \beta_0)}),
% $$
% where $\alpha_{a_k} = \frac{\beta_0 w_{a_k}}{1- w_{a_k}}$ and $w_a = \theta_*^{\top}x_{a}$.

%By the expression of $f$, it is a function of the expected reward $w$. Next, because $B(\tau; \alpha, \beta_0), B(\alpha, \beta_0)$ are decreasing with respect to $\alpha$ and $\alpha$ is in increasing with respect to $w$, then Eq.~\eqref{eq:betaTau} is increasing with respect to $w$. Therefore, the expected reward has the property of monotonicity. For the Lipschitz continuity, by the Lipschitz continuity of expected reward function with conjunctive objective, defined in Eq.~\eqref{eq:AndRewardFunc}, and the continuity of Beta function, we can get $f$ also satisfies the Lipschitz continuity.

%Thus this expected reward function satisfies our setting but can not be fit into previous works, that is, can not be represented by the reward functions with disjunctive/conjunctive objectives. Similarly, there exists a lot of such kind of reward functions, where the real reward is obtained by a second process of observed weight.

% network latency

% If the latency of every edge $e$ follows exponential distribution with mean of $\theta_*^{\top}x_{t,e}$. We call the latency is {\it high} if it is larger than a threshold $\tau$. A path is blocked if some of its edges has high latency. So a path is unblocked with probability
% $$
% \prod_{k=1}^{|A|} (1 - \exp(-(\theta_*^{\top}x_{t,a_k^t})\tau).
% $$

% are observed under the disjunctive objective (resp., conjunctive objective) when under the context of time $t$, that is

% To state our main theorem, we need some definitions. Let
% $$
%   f_{\wedge}^* = \min_{1 \leq t \leq T} f_{\wedge, t}^{\ast}
% $$
% denote the best expected reward of all time under the conjunctive objective.
% Let $p_{t, A}$ (resp., $p_{\wedge, t, A}$) be the {\em probability of full observation of $A$}, that is the probability that all base arms of $A = (a_1, \ldots, a_{\abs{A}})$ are observed under the disjunctive objective (resp., conjunctive objective) when under the context of time $t$, that is 
% \begin{align*}
%   &p_{t, A} = \prod_{k=1}^{\abs{A}-1} (1 - \theta_{\ast}^{\top} x_{t, a_k^t}),\\
%   &p_{\wedge, t, A} = \prod_{k=1}^{\abs{A}-1} (\theta_{\ast}^{\top} x_{t, a_k^t}).
% \end{align*}
% We also define $p^* = \min_{1 \leq t \leq T} \min_{A \in \cS} ~ p_{t, A}$ to be the minimal probability over all time that an action could have all base arms observed under the disjunctive objective. The following is the main theorem on the regret achieved by our C$^3$-UCB algorithm, for both the disjunctive and conjunctive objectives.
% \begin{theorem} %main theorem
%   \label{thm:main}
%   Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_{K} \geq 0$ and $\gamma_k' = 1 - \gamma_k, \gamma'_K \leq \frac{1}{4}f_{\wedge}^{\ast}$. 
%   Let regularization parameter $\lambda$ satisfy $\lambda \geq C_\gamma$, where $C_\gamma = \sum_{k=1}^{K} \gamma_k^2$. 
%   Then for any $\delta > 0$, with probability at least $1 - \delta$, the cumulative expected regrets of our algorithm, C$^3$-UCB for both the conjunctive and disjunctive objectives, satisfy
%   \begin{align}
%     R(T) &\le \frac{\sqrt{2}}{p^*} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))}  \nonumber \\
%     &\qquad \cdot \left(\sqrt{d\ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^{2/d}}\right) } + \sqrt{\lambda}\right) \nonumber \\
%     &=O\left(\frac{d}{p^*} \sqrt{TK} \ln \left(\frac{C_\gamma T}{\delta^{2/d}}\right) \right),
%   \end{align}
%   \begin{align}
%     R_{\wedge}(T) &\le \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \sqrt{TKd\ln(1 + C_\gamma T/(\lambda d))} \nonumber \\
%     &\qquad \cdot \left(\sqrt{d\ln \left( \frac{1 + C_\gamma T/(\lambda d)}{\delta^{2/d}}\right) } + \sqrt{\lambda}\right) \nonumber \\
%     &=O\left(\frac{d}{f_{\wedge}^{\ast}}\sqrt{TK} \ln \left(\frac{C_\gamma T}{\delta^{2/d}}\right) \right).
%   \end{align}
% \end{theorem}

% \subsection{Lower Bounds}

% Suppose we have such a graph: 
% $$
%   \xymatrix{
%   	S_0 \ar[r]^{\mu_{11}} \ar[d]_{\mu_{21}} & B_1 \ar[d]^{\mu_{12}} &\\
%   	B_2 \ar[r]_{\mu_{22}} & S_1
%   }
% $$
% There are two paths in this graph: $A_1 = \{S_0 \to B_1 \to S_1\}$ and $A_2 = \{S_0 \to B_2 \to S_1\}$. 
% Let the feasible action set be $\cS = \{A_1, A_2\}.$ The $\mu$'s on the edges denote the probability of that edge being unblocked. 
% For example, the probability of $S_0 \to B_2$ being unblocked is $\mu_3$. 
% When we select a path, we get reward $1$ if and only if the path is unblocked and get reward $0$ if the path is blocked somewhere. 
% Suppose we have $\mu_3 \geq \mu_1 \geq \mu_1\mu_2 > \mu_3 \mu_4$.
% So the optimal action is $A_1$ with expected reward $\mu_1 \mu_2$ and $A_2$ is suboptimal with expected reward $\mu_3 \mu_4$.

% Similar to the proof of Lai and Robbins 1985, we could get
% $$
% 	\EE[N_{A_2}]\mu_{21} \leq \frac{c \log(T)}{kl(\mu_{22}, \frac{\mu_{11}\mu_{12}}{\mu_{21}})}
% $$

%Next we analyze those times when the learning agent selects path $A_2$ instead of the optimal $A_1$. 
%Notice that if the observed edges are $S_0 \to C \to D$ or $S_0 \to C$, we will not abandon $A_2$ because the expected rewards are $0.9 \times 0.8$ and $0.9$, higher than $0.64$. 
%Only the times when all edges in $A_2$ being observed will help to recognize $A_2$ and abandon it. 
%Recall that the probability of all edges of $A_2$ being observed is $p_{A_2} = 0.9 \times 0.8 = 0.72.$ 
%So we have
% $$
%   \EE[N_{A_2}]p_{A_2} \leq \frac{c \log(T)}{\Delta^2},
% $$
%where $N_{A_2}$ is the number of times that learning agent selects $A_2$ and $\Delta$ is the difference of the best expected reward minus the expected reward of $A_2$.
%Therefore,
% $$
%   \EE[N_{A_2}] \leq \frac{c \log(T)}{p_{A_2} \Delta^2}
% $$
% and
% $$
%   R(T) \leq \EE[\Delta N_{A_2}] \leq \frac{c \log(T)}{p_{A_2} \Delta}.
% $$

% experiment - synthetic - results
\begin{figure*}
	\centering
	\subfigure[Disjunctive, $\gamma=1$]{
		\includegraphics[height = 3 cm, width = 4 cm]{synD-T5-L100-K4-1.eps}}
	\subfigure[Disjunctive, $\gamma=0.9$]{
		\includegraphics[height = 3 cm, width = 4 cm]{synD-T5-L100-K4-9.eps}}
	\subfigure[Conjunctive, $\gamma=1$ ]{
		\includegraphics[height = 3 cm, width = 4 cm]{synC-T5-L100-K4-1.eps}}
	\subfigure[Conjunctive, $\gamma=0.9$ ]{
		\includegraphics[height = 3 cm, width = 4 cm]{synC-T5-L100-K4-9.eps}}
	\caption{Synthetic Data Set, L=100, K=4, d=10}
	\label{fig:synthetic} % 用于交叉引用
\end{figure*}

\begin{figure}
	\centering
	\subfigure[$\gamma = 1$]{
		\includegraphics[height = 3cm, width = 3.95 cm]{movieT30000-L200-K4-1.eps}}
	\subfigure[$\gamma = 0.9$]{
		\includegraphics[height = 3cm, width = 3.95 cm]{movieT30000-L200-K4-9.eps}}
	\caption{MovieLens, L=200, K=4, d=5}
	\label{fig:movielens} % 用于交叉引用
\end{figure}

\begin{figure*}
	\centering
	\subfigure[Network 1239, $\gamma=1$]{
		\includegraphics[height = 3 cm, width = 4 cm]{isp1239-T5-k100-1.eps}}
	\subfigure[Network 1755]{
		\includegraphics[height = 3 cm, width = 4 cm]{network3-1.eps}}
	\subfigure[Network 3257 ]{
		\includegraphics[height = 3 cm, width = 4 cm]{network4-1.eps}}
	\subfigure[Comparison of different $\gamma$ ]{
		\includegraphics[height = 3 cm, width = 4 cm]{network-gamma-with-point.eps}}
	\caption{Network, $d=10$}
	\label{fig:network} % 用于交叉引用
\end{figure*}

\section{Experiments}

We evaluate $C^3$-UCB in three experiments. In the following experiments, we let the position discounts $\gamma_k, k\leq K$ have the form of $\gamma^{k-1}$ for some $\gamma$.

\subsection{Synthetic Data}
\label{sec:ExpSyn}

In the first experiment, we compare $C^3$-UCB to CombCascade on synthetic problems. This problem is a contextual cascading bandit with $L=100$ items and $K=4$, where at each time $t$ the agent recommends $K$ items to the user. At first, we randomly choose $\theta \in \RR^{d-1}$ with $\|\theta \|_2 = 1$ and let $\theta_* = (\frac{\theta}{2}, \frac{1}{2})$. Then at each time $t$, we randomly assign $x'_{t,a} \in \RR^{d-1}$ with $\|x_{t,a}\| = 1$ to arm $a$ and use $x_{t,a} = (x'_{t,a}, 1)$ to be the context information for arm $a$. This processing will guarantee the inner product $\theta_*^{\top}x_{t,a} = \frac{\theta^{\top}x + 1}{2} \in [0,1]$. Next we generate the weight for arm $a$ at time $t$ by a random sample from Bernoulli distribution with mean $\theta_*^{\top}x_{t,a}$.

We experiment on four settings. Under each setting, the learning agent chooses a set of $K$ items out of $L$ ground items. The first two are of the disjunctive objective where the learning agent observes from the first one of the chosen $K$ items until first one with weight $1$; the last two are of conjunctive objective where the learning agent observes from the first item until the first one with weight $0$. Notice that with $\gamma \leq 1$, the oracle to select a feasible action is to select the set of $K$ items with highest UCBs; $\gamma < 1$ will force the items in the optimal action list in the decreasing order of their UCBs and we also use this order when $\gamma = 1$. The results are shown in Fig.~\ref{fig:synthetic} and our algorithm outperforms {\it CombCascade} algorithm because they do not make use of the context information.

\subsection{MovieLens Data}

In this experiment, we evaluate $C^3$-UCB on a problem of contextual recommendations. We experiment on {\it MovieLens} dataset \cite{lam2013movie} of 2015.

The Learning problem is formulated as follows. There is a big sparse matrix $A \in \{0,1\}^{N_1 \times N_2}$ where $A(i,j) = 1$ denotes user $i$ has watched movie $j$. We split $A$ as $H + F$ by putting entry-$1$ in $A$ to $H$ and $F$ with probability $\sim \mathrm{Ber}(p)$ for some fixed $p$. We can regard $H$ as known information about history 'what users have watched' and regard $F$ as future criterion. We use $H$ to derive feature vectors of both users and movies by SVD decomposition, $H = USM^{\top}$ where $U = (u_1; ...;u_{N_1})$ and $M = (m_1;...;m_{N_2})$. At every time $t$, use $x_{t,i} = u_{i}m_j^{\top}$ as the context information of base arm $i$ and randomly choose a user $I_t$. And use $w_t(m) = F(I_t,m)$ as the weight of movie $m$ at time $t$.

In this problem, any combination of $K$ movies is a feasible action. Notice that it might have problem if we use $(u_{I_t}, m_j)$, the concatenation of $u_{I_t}$ and $m_j$, as the context information. Denote $\theta$ by $(\theta_u, \theta_m)$. So the expected mean of a base arm is of the form $\theta_u^{\top} u_{I_t} + \theta_m^{\top} m_j$ but the first term $\theta_u^{\top} u_{I_t}$ is fixed at time $t$ no matter what $\theta_u$ is. To recommend a list of movies by our algorithm, it is equivalent to recommend the first $K$ movies with highest UCBs listing in decreasing order. Thus, it is equivalent to rank the scores $\theta_m^{\top} m_j$ for each movie $j$. So to make use of the user's context information, some quadratic terms of user's information $u_i$ and movie's information $m_j$ are needed. We use the convolution $u_{i}m_j^{\top}$ as the context information of movie $j$ at time $t$.

First, we randomly choose $L=200$ movies and recommend $K=4$ movies on each round. We experiment with $\gamma=1$ (no position discount) and $\gamma=0.9$, and compare our algorithm with {\it CombCascade}. The results are shown in Fig.~\ref{fig:movielens} and demonstrate the advantage to involve context information in real applications.

\subsection{Network Routing Data}

In the last experiment, we evaluate $C^3$-UCB on a problem of network routing with the {\it RocketFuel} dataset \cite{spring2004measuring}.

The ground set $E$ is the set of links in the network and the feasible action set $\cS$ is the set of all paths in the network. Before learning, the environment randomly chooses a $d$-dimensional vector $\theta_* \in [0,1]^d$ and at each time $t$, any link $e$ is assigned with a random $d$-dimensional context vector $x_{t,e}$. Notice here both $\theta$ and $x$ have been processed like section \ref{sec:ExpSyn} so that $\theta_*^{\top}x \in [0,1]$. The weight for each edge $e$ is a sample from Bernoulli distribution with mean $\theta_*^{\top}x$.

At each time $t$, a pair of source and terminal nodes are chosen randomly. The learning agent recommends a feasible path $A = (a_1,...,a_{|A|})$ between $v_s$ and $v_t$ to maximize the expected reward in the conjunctive objective. We experiment with different position discounts and different networks. The results are shown in Fig.~\ref{fig:network}.

We also experiment on the influence of different $\gamma$'s to the regret. In recommendations, the position discounts represent users' satisfaction and the learning agent may not know its true value. Suppose the true position discount $\gamma = 0.9$ and the learning agent sets $\gamma$ by her experience. We experiment on different setting $\gamma \in \{0.8, 0.82, \ldots, 0.98, 1.0\}$ with the true criterion $\gamma = 0.9$. The regrets with different $\gamma$'s are illustrated in Figure~\ref{fig:network}(d). The setting $\gamma$ is more close to the true $\gamma$, the regret is tending to be smaller. This experiment shows when there are position discounts in real application, it is better to represent this information to the model.

% we randomly generate a pair of starting and ending nodes, and the learning agent chooses a routing path between the two nodes to maximize the expected reward with position discount.

%  and the corresponding latency follows exponential distribution with mean $1-\theta^{\top}x$, which has been scaled into $[0,1]$. We call the latency is {\it high} if it is larger than a threshold $\tau$. 
% $$
% \prod_{k=1}^{|A|} (1- \exp(-\frac{\tau}{1-\theta^{\top}x_{t,a_k}})).
% $$
% Notice that the expression is not of the form $\prod_{k=1}^{|A|}\theta_*^{\top}x_{t,a_k}$ in our model. However, our algorithm could also apply to it.

% the environment randomly choose

\section{Conclusions}

In this paper, we propose contextual combinatorial cascading bandits with position discounts, where each action is an ordered list and only a prefix of the action can be observed each time. We propose a C$^3$-UCB algorithm and prove a cumulative regret bound for general reward functions and two special reward functions. The experiments conducted demonstrate the advantage to involve context information and position discounts. In the next, we want to derive a lower bound for this problem.

% In the unusual situation where you want a paper to appear in the
% references without citing it in the main text, use \nocite
%\nocite{langley00}

\bibliography{cascade_reference}
\bibliographystyle{icml2016}

\appendix

\section{Proof of Theorem~\ref{thm:main}}

The following lemma provides the important result on the regret at each time $t$ in terms of feature vectors of base arms at the time.

\begin{lemma} %lem:DeltaEstimate
	\label{lem:DeltaEstimate}
	For any time $t$ and $A = (a_1, \ldots, a_{\abs{A}})$, if $f$ satisfies the two assumptions, the properties of monotonicity and Lipschitz continuity, 
	%and $f(A_t^*, \bU_t) \leq f(A, \bU_t)$, 
	then we have
	$$
		R(t, \bA_t) \leq B \sum_{k=1}^{\abs{\bA_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t, \ba_k}}_{\bV_{t-1}^{-1}}.
	$$
\end{lemma}
\begin{proof}
	By Lemma \ref{lem:estimateU}, $w_t \leq \bU_t$. Then
	\begin{align*}
		f(\bA_t, \bU_t) \geq \alpha f(\bA_t^{\bU_t}, \bU_t) \geq \alpha f(A_t^*, \bU_t) \geq \alpha f(A_t^*, w_t),
	\end{align*}
	where $f(\bA_t^{\bU_t}, \bU_t) = \max_{A \in \cS} f(A, \bU_t)$. The first inequality is because we choose $\bA_t$ by the $\alpha$-approximation oracle with input $\bU_t$; the second inequality is because the maximum of $f(\bA_t^{\bU_t}, \bU_t)$ given $\bU_t$; the third inequality is by the monotonicity of $f$ and the property $\bU_t \geq w_t$. Then
	\begin{align*}
		R(t, \bA_t) = &\alpha f(A_t^*, w_t) - f(\bA_t, w_t) \\
		\leq &f(\bA_t, \bU_t) - f(\bA_t, w_t).
	\end{align*}
	By Lipschitz continuity of $f$ and the definition of $\bU_t$ in Eq.\eqref{eq:defU},
	$$
	f(\bA_t, \bU_t) \leq f(\bA_t, \theta_{\ast}^{\top}x_t) + B \sum_{k=1}^{\abs{\bA_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t, \ba_k^t}}_{\bV_{t-1}^{-1}}.
	$$
	Then we have
	\begin{align*}
		R(t, A) \leq &f(\bA_t, \bU_t) - f(\bA_t, w_t) \\
		\leq &B \sum_{k=1}^{\abs{\bA_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t, \ba_k^t}}_{\bV_{t-1}^{-1}}.
	\end{align*}
\end{proof}

Notice that the upper bound of $R(t, \bA_t)$ is in terms of all base arms of $\bA_t$. However, it is hard to estimate an upper bound for $\sum_{t=1}^T \sum_{k=1}^{\abs{\bA_t}} \norm{ x_{t, \ba_k^t} }_{ \bV_{t-1}^{-1} }$ because $\bV_t$ only contains information of observed base arms. Thus we need the following lemma.

\begin{lemma} %lem:DeltaEsimateWithP*
	\label{lem:DeltaEsimateWithP*}
	Suppose the equation (\ref{eq:estimateTheta}) holds for time $t-1$. Then
	$$
	\EE_t[R(t, \bA_t)] \leq \frac{B}{p^*} \EE_t \left[ \bbeta_{t-1}(\delta) \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
	$$
\end{lemma}
\begin{proof}
	\begin{align}
		&\EE[R(t, \bA_t)|\cH_t]  \nonumber \\
		=& \EE \left[ \left. R(t, \bA_t) \EE \left[ \left. \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t = \abs{\bA_t}\} \right| \bA_t \right]  \right| \cH_t \right] \label{eq:addOt}\\
		=& \EE \left[ \left. R(t, \bA_t) \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t = \abs{\bA_t}\}  \right| \cH_t \right] \label{eq:removeE}\\
		\leq &\frac{1}{p^{\ast}} \EE \left[ \left. R(t, \bA_t) \bOne\{\bO_t = \abs{\bA_t}\}  \right| \cH_t \right] \label{eq:relaxpstar}\\
		\leq &\frac{B}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\abs{\bA_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \bOne\{\bO_t = \abs{\bA_t}\}  \right| \cH_t \right] \label{eq:useradiusbound}\\
		\leq &\frac{B}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\bO_t} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \right| \cH_t \right]. \label{eq:relaxindicator}
	\end{align}
	Equality \eqref{eq:addOt} is because when $\bA_t$ is fixed, $p_{t, \bA_t}$ is the probability of $\bO_t = \abs{\bA_t}$, and thus the inner expectation is $1$. Equality \eqref{eq:removeE} is because when the history $\cH_t$ is fixed, $\bA_t$ is fixed by the deterministic algorithm C$^3$-UCB, and thus there is no need to write the conditional expectation. Inequality~\eqref{eq:relaxpstar} is because $p_{t,\bA_t} \geq p^*$ by definition. Inequality~\eqref{eq:useradiusbound} is by Lemma \ref{lem:DeltaEstimate} and the fact $f(A_t^*, \bU_t) \leq f(\bA_t, \bU_t)$, which is true because algorithm C$^3$-UCB selects action $\bA_t$ as the best action with respect to $\bU_t$. Inequality~\eqref{eq:relaxindicator} is by simply arguing on the two cases for the indicator function $\bOne\{\bO_t = \abs{\bA_t}\}$.
\end{proof}

% \begin{lemma} %lem:DeltaEstimate
%   \label{lem:DeltaEstimate}
%   For any time $t$ and $A = (a_1, \ldots, a_{\abs{A}})$, if $f(A_t^*, \bU_t) \leq f(A, \bU_t)$, then we have
%   $$
%     R(t,A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}};
%   $$
%   if $f_{\wedge}(A_t^*, \bU_t) \leq f_{\wedge}(A, \bU_t)$, then we have
%   $$
%     R_{\wedge}(t, A) \leq \sum_{k=1}^{\abs{A}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}}.
%   $$
% \end{lemma}
% \begin{proof}
%   By Lemma \ref{lem:estimateU}, $\theta_{\ast}^{\top}x_t \leq \bU_t$. Then by Lemma \ref{lem:increasing},
%   $$
%     f_t^{\ast} = f(A_t^{\ast}, \theta_{\ast}^{\top}x_t) \leq f(A_t^{\ast}, \bU_t).
%   $$
%   By Lemma \ref{lem:estimateTech} and the definition of $\bU_t$ in Eq.\eqref{eq:defU},
%   $$
%     f(A, \bU_t) \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
%   $$
%   By our condition that $f(A_t^{\ast}, \bU_t) \leq f(A, \bU_t)$, we have 
%   $$
%     f_t^{\ast} \leq f(A, \theta_{\ast}^{\top}x_t) + \sum_{k=1}^{\abs{A}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
%   $$
%   Therefore,
%   $$
%     R(t, A) = f_t^{\ast} - f(A, \theta_{\ast}^{\top}x_t) \leq \sum_{k=1}^{\abs{A}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t, a_k^t}}_{\bV_{t-1}^{-1}}.
%   $$
%   The result for the conjunctive case can be deduced in a similar way.
% \end{proof}

% Notice that if we substitute $A$ by $\bA_t$ in Lemma \ref{lem:DeltaEstimate}, then the upper bound of $R(t, \bA_t)$ is in terms of all base arms of $\bA_t$. 
% However, it is hard to estimate an upper bound for $\sum_{t=1}^T \sum_{k=1}^{\abs{\bA_t}} \norm{ x_{t, \ba_k^t} }_{ \bV_{t-1}^{-1} }$ because $\bV_t$ only contains observed base arms. Thus we need the following lemma.

% \begin{lemma} %lem:DeltaEsimateWithP*
%   \label{lem:DeltaEsimateWithP*}
%   Let $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$ and $\gamma_k' = 1 - \gamma_k$.
%   Suppose the equation (\ref{eq:estimateTheta}) holds for time $t-1$. Then
%   $$
%     \EE_t[R(t, \bA_t)] \leq \frac{1}{p^*} \EE_t \left[ \bbeta_{t-1}(\delta) \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
%   $$
%   If we have
%   $$
%     \gamma_K' \le \frac{1}{4} f_{\wedge}^{\ast} = \frac{1}{4} \min_{1 \leq t \leq T} f_{\wedge, t}^{\ast},
%   $$
%   then
%   $$
%     \EE_t [R_{\wedge}(t, \bA_t) ] \leq \frac{8}{f_{\wedge}^{\ast}} \EE_t \left[ \bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
%   $$
% \end{lemma}
% \begin{proof}
%   For the disjunctive case,  
%   \begin{align}
%     &\EE[R(t, \bA_t)|\cH_t]  \nonumber \\
%     =& \EE \left[ \left. R(t, \bA_t) \EE \left[ \left. \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t\geq\abs{\bA_t}\} \right| \bA_t \right]  \right| \cH_t \right] \label{eq:addOt}\\
%     =& \EE \left[ \left. R(t, \bA_t) \frac{1}{p_{t, \bA_t}} \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right] \label{eq:removeE}\\
%     \leq &\frac{1}{p^{\ast}} \EE \left[ \left. R(t, \bA_t) \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right] \label{eq:relaxpstar}\\
%     \leq &\frac{1}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\abs{\bA_t}} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \bOne\{\bO_t\geq\abs{\bA_t}\}  \right| \cH_t \right] \label{eq:useradiusbound}\\
%     \leq &\frac{1}{p^{\ast}} \EE \left[ \left. \sum_{k=1}^{\bO_t} \gamma_k \bbeta_{t-1}(\delta)\norm{x_{t,a_k}}_{\bV_{t-1}^{-1}} \right| \cH_t \right]. \label{eq:relaxindicator}
%   \end{align}
%   Equality \eqref{eq:addOt} is because when $\bA_t$ is fixed, $p_{t, \bA_t}$ is the probability of $\bO_t\geq\abs{\bA_t}$, and thus the inner expectation is $1$. 
%   Equality \eqref{eq:removeE} is because when the history $\cH_t$ is fixed, $\bA_t$ is fixed by the deterministic algorithm C$^3$-UCB, and thus there is no need to write the conditional expectation. 
%   Inequality~\eqref{eq:relaxpstar} is because $p_{t,\bA_t} \geq p^*$ by definition. 
%   Inequality~\eqref{eq:useradiusbound} is by Lemma \ref{lem:DeltaEstimate} and the fact $f(A_t^*, \bU_t) \leq f(\bA_t, \bU_t)$, which is true because algorithm C$^3$-UCB selects action $\bA_t$ as the best action with respect to $\bU_t$. 
%   Inequality~\eqref{eq:relaxindicator} is by simply arguing on the two cases for the indicator function $\bOne\{\bO_t\geq\abs{\bA_t}\}$.

%   For the conjunctive case, by Lemma~\ref{lem:prefixexists} there exists a prefix $\bB_t$ of $\bA_t$ such that $p_{\wedge, t, \bB_t} \geq \frac{1}{2}f_{\wedge, t}^* - \gamma'_{|\bB_t|}$ and $R_{\wedge}(t, \bB_t) \geq \frac{1}{2} R_{\wedge}(t, \bA_t)$. 
%   By the assumption that $\gamma'_{|\bB_t|}\le \gamma'_K \le \frac{1}{4} f^*_{\wedge} \le \frac{1}{4} f^*_{\wedge,t}  $, we have $p_{\wedge, t, \bB_t} \geq \frac{1}{4}f_{\wedge, t}^*$. 
%   Then, similar to the derivation from Eq.\eqref{eq:addOt} to Eq.\eqref{eq:relaxpstar}, we have
%   \begin{align}
%     &\EE_t[R_{\wedge}(t, \bA_t)] \leq 2 \EE[R_{\wedge}(t, \bB_t) \left. \right| \cH_t] \nonumber \\
%     =& 2 \EE \left[ \left. R_{\wedge}(t, \bB_t) \EE \left[ \left. \frac{1}{p_{\wedge, t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\} \right| \bB_t \right]  \right| \cH_t \right] \nonumber \\
%     =& 2\EE \left[ \left. R_{\wedge}(t, \bB_t) \frac{1}{p_{\wedge, t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right] \nonumber \\
%     \leq &\frac{8}{f_{\wedge, t}^{\ast}} \EE \left[ \left. R_{\wedge}(t, \bB_t) \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right]. \label{eq:btOneotbt}
%   \end{align}
%   Next, we have
%   \begin{equation} \label{eq:andstarbt}
%     f_{\wedge}(A_t^*, \theta_*^{\top}x_t) \leq f_{\wedge}(A_t^*,\bU_t) \leq f_{\wedge}(\bA_t,\bU_t) \leq f_{\wedge}(\bB_t,\bU_t),
%   \end{equation}
%   where the first inequality is by Lemmas \ref{lem:estimateU} and \ref{lem:increasing}, the second inequality is by the C$^3$-UCB algorithm, in which the best action $\bA_t$ is selected with respect to $\bU_t$, and the last inequality is by Lemma~\ref{lem:prefixRelation}~(2).

%   Then by Lemma \ref{lem:estimateTech},
%   \begin{equation} \label{eq:btuttrue}
%     f_{\wedge}(\bB_t,\bU_t) \leq f_{\wedge}(\bB_t, \theta_*^{\top}x_t) + \sum_{k=1}^{\abs{\bB_t}}\gamma_k\bbeta_{t-1}(\delta)\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}.
%   \end{equation}
%   Therefore, combining Eq.\eqref{eq:andstarbt} and \eqref{eq:btuttrue}, we have
%   \begin{align}
%     R_{\wedge}(t, \bB_t) & = f_{\wedge, t}^{\ast} - f_{\wedge}(\bB_t, \theta_{\ast}^{\top}x_t) \nonumber \\
%     & \leq \sum_{k=1}^{\abs{\bB_t}}\bbeta_{t-1}(\delta)\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}. \label{eq:sumbetax}
%   \end{align}
%   Finally from Inequalities~\eqref{eq:btOneotbt} and \eqref{eq:sumbetax}， we can derive the result
%   \begin{align*}
%     &\EE_t[R_{\wedge}(t, \bA_t)] \\
%     \leq& \frac{8}{f_{\wedge, t}^*} \EE \left[\bOne\{ \bO_t \geq \abs{\bB_t} \}\sum_{k=1}^{\abs{\bB_t}}\bbeta_{t-1}(\delta)\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right]\\
%     \leq& \frac{8}{f_{\wedge}^*} \EE \left[ \sum_{k=1}^{ \bO_t } \bbeta_{t-1}(\delta) \norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
%   \end{align*}
% \end{proof}

From the above lemma, we can bound the regret at time $t$ in terms of the observed base arms, for which we further bound in Lemma~\ref{lem:SumXiEstimateInDet}.
But before that, we need the following technical lemma.

\begin{lemma} %lem:detTech
  \label{lem:detTech}
  Let $x_i \in \RR^{d \times 1}, 1 \leq i \leq n$. Then we have
  $$
    \det\left(I + \sum_{i=1}^n x_i x_i^{\top}\right) \geq 1 + \sum_{i=1}^n \norm{x_i}_2^2.
  $$
\end{lemma}
\begin{proof}
  Denote the eigenvalues of $I + \sum_{i=1}^n x_i x_i^{\top}$ by $1+\alpha_1,...,1+\alpha_d$ with $\alpha_j \geq 0$, $1\leq j\leq d$. Then
  \begin{align*}
    &\det(I + \sum_{i=1}^n x_i x_i^{\top})= \prod_{j=1}^d (1 + \alpha_j)\\
    &\geq 1 +\sum_{j=1}^d \alpha_j =1-d + \sum_{i=1}^d (1+\alpha_i) \\
    &=1-d + \trace(I + \sum_{i=1}^n x_i x_i^{\top})= 1-d + d + \sum_{i=1}^n \norm{x_i}_2^2\\
    &=1 + \sum_{i=1}^n \norm{x_i}_2^2.
  \end{align*}
\end{proof}

\begin{lemma} % lem:SumXiEstimateInDet
  \label{lem:SumXiEstimateInDet}
  \CLemmaSumXiEstimateInDet
\end{lemma}
\begin{proof}
	\begin{align}
	&\det(\bV_t) = \det\left(\bV_{t-1} + \sum_{k=1}^{\bO_t} (\gamma_k x_{t,\ba_k^{t}})(\gamma_k x_{t, \ba_k^{t}}^{\top}) \right) \nonumber \\
	&=\det(\bV_{t-1}) \notag \\
	&~~~\cdot \det \left(I + \sum_{k=1}^{\bO_t} \gamma_k \bV_{t-1}^{-1/2}x_{t,\ba_{k}^{t}} (\gamma_k \bV_{t-1}^{-1/2}x_{t,\ba_{k}^{t}})^{\top} \right) \label{eq:Vfactor}\\
	&\geq \det(\bV_{t-1}) \left(1 + \sum_{k=1}^{\bO_t} \norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2 \right) \label{eq:detGeq} \\
	&\geq \det(\lambda I)\prod_{s=1}^{t} \left(1 + \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \right), \label{eq:repeatV}
	\end{align}
	where Equality~\eqref{eq:Vfactor} is by the fact that $V+U = V^{1/2} (I + V^{-1/2} U V^{-1/2}) V^{1/2}$ for a symmetric positive definite matrix $V$, Inequality \eqref{eq:detGeq} is due to Lemma \ref{lem:detTech}, and Inequality~\eqref{eq:repeatV} is by repeatedly applying Inequality \eqref{eq:detGeq}.

	Because
	$$
	\norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \leq \norm{\gamma_k x_{s,\ba_k^s}}_2^2/\lambda_{\min}(\bV_{s-1}) \leq \gamma_k^2 /\lambda,
	$$
	where $\lambda_{\min}(\bV_{s-1})$ is the minimum eigenvalue of $\bV_{s-1}$,  we have 
	$$
	\sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \leq \frac{1}{\lambda} \sum_{k=1}^{K} \gamma_k^2 = C_\gamma /\lambda \leq 1
	$$
	Using the fact that $ 2\ln(1+u) \geq u$ for any $u \in [0,1]$, we get
	\begin{align*}
	&\sum_{s=1}^t \sum_{k=1}^{\bO_s}\norm{\gamma_k x_{s,\ba_{k}^s}}_{\bV_{s-1}^{-1}}^2 \\
	&\leq 2\sum_{s=1}^t\ln \left(1 + \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \right)\\
	& = 2 \ln \prod_{s=1}^{t} \left(1 + \sum_{k=1}^{\bO_s} \norm{\gamma_k x_{s,\ba_k^s}}_{\bV_{s-1}^{-1}}^2 \right) \\
	&\le 2 \ln \left(\frac{\det(\bV_t)}{\det(\lambda I)} \right),
	\end{align*}
	where the last inequality is from Inequality~\eqref{eq:repeatV}.
\end{proof}

Finally the last lemma bounds $\det(\bV_t)$.

\begin{lemma} % det(V_t)
	\label{lem:detVt}
	\CLemmaDetVt
\end{lemma}
\begin{proof}
	To prove $\det(\bV_t)$ is increasing with respect to $t$, it is enough to prove that
	$$
	\det(V + xx^{\top}) \geq \det(V),
	$$
	for any symmetric positive definite matrix $V \in \RR^{d \times d}$ and column vector $x \in \RR^{d\times 1}$. In fact,
	\begin{align*}
	&\det(V + xx^{\top}) \\
	=& \det(V) \det(I + V^{-1/2}x x^{\top} V^{-1/2})\\
	=& \det(V) \det(1 + \norm{V^{-1/2}x}^2)\\
	\geq &\det(V).
	\end{align*}
	The second equality above is due to Sylvester's determinant theorem, which states that $\det(I + AB) = \det(I +BA)$.

	Let the eigenvalues of $\bV_t$ be $\lambda_1, \ldots, \lambda_d > 0$. Then
	\begin{align*}
	&\det(\bV_t) = \lambda_1 \cdot \cdots \cdot \lambda_d \\
	\leq & \left( \frac{\lambda_1 + \ldots + \lambda_d}{d} \right)^d = (\trace(\bV_t)/d)^d.
	\end{align*}
	Also,
	\begin{align*}
	&\trace(\bV_t)\\
	& = \trace(\lambda I) + \sum_{s=1}^t \sum_{k=1}^{\bO_s} \gamma_k^2 \trace(x_{s,\ba_k^s} x_{s,\ba_k^s}^{\top})\\	
	& = d \lambda + \sum_{s=1}^t \sum_{k=1}^{\bO_s} \gamma_k^2 \norm{x_{s,\ba_k^s}}_2^2\\
	& \leq d \lambda + \sum_{s=1}^t\sum_{k=1}^{K}\gamma_k^2 = d \lambda + t C_\gamma.
	\end{align*}
	Thus we have $\det(\bV_t) \leq (\lambda + C_\gamma t/d)^d.$
\end{proof}

Now we are ready to prove the main theorem.

\begin{proof}[ of Theorem \ref{thm:main}]
	Suppose inequality (\ref{eq:estimateTheta}) holds for all time $t$, which is true with probability $1-\delta$. Then we have
	\begin{align}
	&R(T) = \EE \left[\sum_{t=1}^{n} \EE_{t}[R(t, \bA_t)] \right] \notag\\
	&\leq \EE\left[\sum_{t=1}^{n} \frac{B}{p^*} \EE_t \left[\bbeta_{t-1}(\delta) \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}\right] \right] \label{eq:pfMdeltaEst}\\
	&\leq \EE\left[\sum_{t=1}^{n} \frac{B}{p^*} \bbeta_{T}(\delta) \EE_t \left[ \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}\right] \right] \label{eq:pfMbeta}\\
	&\leq \EE \left[\frac{B}{p^*} \bbeta_T(\delta) \sqrt{\left(\sum_{t=1}^{n} \bO_t\right) \left( \sum_{t=1}^{n} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2 \right)} \right ]  \label{eq:pfMmeanEq}\\
	&\leq \EE\left[\frac{B}{p^*} \left( R \sqrt{ \ln \left( \frac{\det(\bV_n)}{\lambda^d \delta^2} \right)} + \sqrt{\lambda} \right) \right. \notag\\
	& \left. \qquad \qquad \cdot \sqrt{nK \cdot 2\ln \left(\frac{\det(\bV_n)}{\lambda^d} \right)} \right] \label{eq:pfMsumXi}\\
	&\leq \frac{\sqrt{2}B}{p^*} \left(R\sqrt{ d \ln \left( \frac{1 + C_\gamma n/(\lambda d)}{\delta^{2/d}} \right) } + \sqrt{\lambda} \right) \notag\\
	&\qquad \qquad \cdot \sqrt{nKd\ln(1 + C_\gamma n/(\lambda d))}. \label{eq:pfMdetEst}
	\end{align}
	Inequality \eqref{eq:pfMdeltaEst} is by Lemma \ref{lem:DeltaEsimateWithP*}. 
	Inequality \eqref{eq:pfMbeta} is because $\bbeta_{t}(\delta)$ is increasing with respect to $t$, derived by the definition of $\bbeta_t(\delta)$ (Eq.\eqref{eq:definebeta}) and Lemma \ref{lem:detVt}. 
	Inequality \eqref{eq:pfMmeanEq} is by the mean inequality. 
	Inequality \eqref{eq:pfMsumXi} is because of the definition of $\bbeta_t(\delta)$ (Eq.\eqref{eq:definebeta}) and Lemma \ref{lem:SumXiEstimateInDet}. 
	And the last inequality \eqref{eq:pfMdetEst} is because estimate of $\det(\bV_t)$ by Lemma \ref{lem:detVt}.
\end{proof}

\section{Proof of Corollary~\ref{cor:or}}

In this section, we prove the expected reward function
$$
f(A,w) = \sum_{k = 1}^{\abs{A}} \gamma_{k} \prod_{i=1}^{k-1} (1 - w(a_i)) w(a_k),
$$
where $A = (a_1, \ldots, a_{|A|})$, defined in~Eq.\eqref{eq:OrRewardFunc}, satisfies the properties of monotonicity and Lipschitz continuity.

\begin{lemma}[monotonicity]
	\label{lem:orMonotonicity}
	Suppose $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$. Then $f(A, w)$ is increasing with respect to $w$; that is, if $0 \leq w \leq w' \leq 1$ as column vectors in $\RR^L$, then for any $A \in \prod^{\leq K}$, it holds that
	$$
	f(A, w) \leq f(A, w').
	$$
\end{lemma}
\begin{proof}
	Without the loss of generality, we prove for the case of $A = (1, \ldots, m)$, where $1 \leq m \leq K$. First, for $1 \leq k \leq m$, we have
	\begin{align*}
		&\gamma_{k} - \sum_{i=k+1}^m \gamma_i \prod_{j = k + 1}^{i - 1} (1 - w_j') w_i'\\
		\geq &~\gamma_k [1 - \sum_{i=k+1}^m \prod_{j=k+1}^{i-1}(1 - w_j') w_i']\\
		\geq &~\gamma_{k} \cdot 0 = 0.
	\end{align*}
	Then it implies that
	\begin{align*}
		&\gamma_k w_k + (1 - w_k)\sum_{i=k+1}^m \gamma_i \prod_{j=k+1}^{i-1}(1 - w_j') w_i'\\
		\leq &\gamma_k w_k' + (1 - w_k')\sum_{i=k+1}^m \gamma_i \prod_{j=k+1}^{i-1}(1 - w_j') w_i'.\\
	\end{align*}
	Therefore, 
	\begin{align*}
		& f(A; w_1, \dots, w_k, w_{k+1}', \dots, w_m')\\
		&=\sum_{i=1}^{k-1} \gamma_i \prod_{j=1}^{i-1}(1 - w_j) w_i + \prod_{j=1}^{k-1}(1 - w_j) \\
		&\qquad \cdot [\gamma_k w_k + (1 - w_k)\sum_{i=k+1}^m \gamma_i \prod_{j=k+1}^{i-1}(1 - w_j') w_i']\\
		&\leq \sum_{i=1}^{k-1} \gamma_i \prod_{j=1}^{i-1}(1 - w_j) w_i + \prod_{j=1}^{k-1}(1 - w_j) \\
		&\qquad \cdot [\gamma_k w_k' + (1 - w_k')\sum_{i=k+1}^m \gamma_i \prod_{j=k+1}^{i-1}(1 - w_j') w_i']\\
		&=f(A; w_1, \ldots, w_{k-1}, w_{k}', \ldots, w_m').
	\end{align*}
\end{proof}

\begin{lemma}[Lipschitz continuity]
	\label{lem:orLip}
	Suppose $0 \leq \gamma_k \leq 1, k \leq K$. Let $w, w'\in [0,1]^{L}$. Then
	$$
	|f(A, w) - f(A, w')| \leq \sum_{k=1}^{|A|} \gamma_k |w(a_k) - w'(a_k)|,
	$$
	where $A = (a_1, \ldots, a_{|A|})$.
\end{lemma}
\begin{proof}
	Without the loss of generality, we prove for the case of $A = (1, \ldots, m)$, where $1 \leq m \leq K$. First, suppose $w \leq w'$ and $w' = w + v$.

	We prove this by induction. It holds obviously for $m = 1$. Suppose it holds for $m$. Then
	\begin{align*}
		&f(A^{(m+1)}, w')\\
		=&f(A^{(m)}, w') + \gamma_{m+1}\prod_{k=1}^m(1 - w'_k) w'_{m+1}\\
		\leq &f(A^{(m)}, w') +  \gamma_{m+1} \prod_{k=1}^m(1 - w_k) (w_{m+1} + v_{m+1})\\
		\leq &f(A^{(m)}, w) + \sum_{k=1}^m \gamma_k v_k \\
		&\qquad + \gamma_{m+1} \prod_{k=1}^m(1 - w_k) w_{m+1} + \gamma_{m+1} v_{m+1}\\
		=& f(A^{(m+1)}, w) + \sum_{k=1}^{m+1} \gamma_k v_k.
	\end{align*}
	So it holds for $m+1$.

	For general $w,w' \in [0,1]^L$, by last lemma, we have
	$$
	f(A, w\wedge w') \leq f(A, w), f(A, w') \leq f(A, w\vee w'),
	$$
	where 
	$$
	(w\wedge w')_k = \min\{w_k, w'_k\}, ~ (w\vee w')_k = \max\{w_k, w'_k\}.
	$$
	Then
	\begin{align*}
		&|f(A,w) - f(A,w')| \leq f(A, w\vee w') - f(A, w\wedge w')\\
		\leq & \sum_{k=1}^{|A|} \gamma_k [(w\vee w')(a_k) - (w\wedge w')(a_k)]\\
		=& \sum_{k=1}^{|A|} \gamma_k |w(a_k) - w'(a_k)|
	\end{align*}
\end{proof}

We have proved the monotonicity and Lipschitz continuity of expected reward function $f$. Then by Theorem \ref{thm:main}, the corollary can be obtained.

\section{Proof of Theorem~\ref{thm:and}}

In this section, we prove Theorem \ref{thm:and}. Recall the expected reward function is
$$
f(A,w) = \sum_{k = 1}^{\abs{A}} (1 - \gamma_k) \prod_{i = 1}^{k - 1} w(a_i)(1 - w(a_k)) + \prod_{i=1}^{\abs{A}}w(a_i),
$$
where $A = (a_1, \ldots, a_{|A|})$, as defined in~Eq.\eqref{eq:AndRewardFunc}. First, similar to last section, we prove this reward function satisfies the properties of monotonicity and Lipschitz continuity.

\begin{lemma}[monotonicity]
	\label{lem:andMonotonicity}
	Suppose $1 \geq \gamma_1 \geq \cdots \geq \gamma_K \geq 0$. Then $f(A, w)$ is increasing with respect to $w$; that is, if $0 \leq w \leq w' \leq 1$ as column vectors in $\RR^L$, then for any $A \in \prod^{\leq K}$, it holds that
	$$
	f(A, w) \leq f(A, w').
	$$
\end{lemma}
\begin{proof}
	Denote $A = (a_1, \ldots, a_{|A|})$. Because
	\begin{align*}
	&1 - f(A, 1-w) \\
	=& 1- \sum_{k = 1}^{\abs{A}} (1 - \gamma_k) \prod_{i = 1}^{k - 1} (1-w(a_i))w(a_k) - \prod_{i=1}^{\abs{A}}(1-w(a_i))\\
	=& \sum_{k = 1}^{\abs{A}} \gamma_{k} \prod_{i=1}^{k-1} (1 - w(a_i)) w(a_k),
	\end{align*}
	then by the proof of Lemma \ref{lem:orMonotonicity} and $1-f(A,1-w)$ is increasing in $w$, we can get $f$ is increasing in $w$.
\end{proof}

\begin{lemma}[Lipschitz continuity]
	\label{lem:andLip}
	Suppose $0 \leq \gamma_k \leq 1, k \leq K$. Let $w, w'\in [0,1]^{L}$. Then
	$$
	|f(A, w) - f(A, w')| \leq \sum_{k=1}^{|A|} \gamma_k |w(a_k) - w'(a_k)|,
	$$
	where $A = (a_1, \ldots, a_{|A|})$.
\end{lemma}
\begin{proof}
	Similar to the proof of Lemma \ref{lem:orLip}. It is enough to prove when $w \leq w', w' = w + v $ and $A = (1, \ldots, m), m \leq K$.

	We prove this by induction. It holds obviously for $m = 1$. Suppose it holds for $m$. Then
	\begin{align*}
		&f(A^{(m+1)}, w') \\
		= &f(A^{(m)}, w') -\prod_{k=1}^{m} w'_k \\
		&\qquad+ (1-\gamma_{m+1}) (\prod_{k=1}^{m} w'_k) (1 - w'_{m+1})+ \prod_{k=1}^{m+1} w'_k\\
		=& f(A^{(m)}, w') - \gamma_{m+1} (\prod_{k=1}^{m} w'_k) (1 - w'_{m+1})\\
		\leq& f(A^{(m)}, w') - \gamma_{m+1} (\prod_{k=1}^{m} w_k) (1 - w'_{m+1})\\
		\leq& f(A^{(m)}, w') -\gamma_{m+1} (\prod_{k=1}^{m} w_k)  (1 - w_{m+1} - v_{m+1})\\
		\leq& (f(A^{(m)}, w) +  \sum_{k=1}^{m} \gamma_k v_k) \\
		&\qquad - \gamma_{m+1}  (\prod_{k=1}^{m} w_k)(1 - w_{m+1}) + \gamma_{m+1} v_{m+1}\\
		\leq& (A^{(m+1)}, w) + \sum_{k=1}^{m+1} \gamma_k v_k.
	\end{align*}
\end{proof}

The next lemma provides some properties about a prefix of action $A$ in the conjunctive case, which leads to the finding of a prefix $B$ of $A$ with similar regret and probability of full observation as those of $A$, as given in Lemma~\ref{lem:prefixexists}.

\begin{lemma} %lem:prefixRelation
	\label{lem:prefixRelation}
	Suppose $1 = \gamma_1 \geq \gamma_2 \geq \cdots \geq \gamma_K \geq 0$. Let $A = (a_1, ..., a_{\abs{A}})$. Let $B_k = (a_1, ..., a_k), k \leq \abs{A}$ be a prefix of $A$. The expected weights are denoted by $w$. The probability of full observation of $A$, $p_{A}$, can be formulated as
	$$
	p_{A} = \prod_{k=1}^{\abs{A}-1} w_{a_k}.
	$$
	Then for the problem with conjunctive objective and $k < \abs{A}$, we have the following properties:
	\begin{itemize}
		\item[(1)] $f(A, w) \leq (1-\gamma_{\abs{A}}) + \gamma_{\abs{A}} p_{A}$;
		\item[(2)] $f(B_{k+1}, w) \leq f(B_k, w)$;
		\item[(3)] $p_{B_{k+1}} \leq p_{B_k}$;
		\item[(4)] $f(B_k, w) \leq (1-\gamma_{k+1}) + \gamma_{k+1} p_{B_{k+1}}$.
	\end{itemize}
\end{lemma}
\begin{proof}
	Let $m = |A|$. By the definition of $f(A,w)$ in Eq.~\eqref{eq:AndRewardFunc}, we have
	\begin{itemize}
		\item[(1)]
		\begin{align*}
		&f(A,w) \\
		=& \sum_{k = 1}^{\abs{A}} (1-\gamma_k) \prod_{i = 1}^{k - 1} w(a_i) (1 - w(a_k)) + \prod_{i=1}^{\abs{A}}w(a_i) \\
		\leq& (1-\gamma_m)(1 - p_{A} w_m) + p_{A} w_m \\
		\leq& (1-\gamma_m) + \gamma_m p_{\wedge, A};
		\end{align*}

		\item[(2)]
		\begin{align*}
		&f(B_k, w) - f(B_{k+1}, w)\\
		=&\prod_{i=1}^{k}w_i - (1-\gamma_{k+1}) (\prod_{i=1}^{k}w_i) (1 - w_{k+1}) \\
		&\qquad \qquad - (\prod_{i=1}^{k}w_i) w_{k+1}\\
		=&\gamma_{k+1} (\prod_{i=1}^{k}w_i) (1 - w_{k+1}) \geq 0;
		\end{align*}

		\item[(3)]
		$p_{B_{k+1}} = \prod_{i=1}^{k} w_i \leq \prod_{i=1}^{k-1} w_i = p_{B_k}$;

		\item[(4)]
		\begin{align*}
		f(B_k, w) & \leq (1-\gamma_k) (1 - p_{B_{k+1}}) + p_{B_{k+1}}\\
		&\leq (1-\gamma_{k+1}) (1 - p_{B_{k+1}}) + p_{B_{k+1}} \\
		&= (1-\gamma_{k+1}) + \gamma_{k+1} p_{B_{k+1}}
		\end{align*}
	\end{itemize}
\end{proof}

\begin{lemma} %lem:prefixexists
	\label{lem:prefixexists}
	\CLemmaPrefixExi
\end{lemma}
\begin{proof}
	$w_t = \theta_*^{\top} x_t$. If $f(A, w_t) \geq \frac{1}{2} f_{t}^{\ast}$, then by Lemma \ref{lem:prefixRelation} (1),
	$$
	p_{t, A} \geq \frac{1}{2} f_{t}^{\ast} - 1 + \gamma_{\abs{A}}.
	$$
	In this case, we set prefix $B = A$.

	Now suppose $f(A, w_t) \leq \frac{1}{2} f_{t}^{\ast}$. Let
	$$
	x_k = f(B_k,w), ~~ y_k = 1 - \gamma_k + \gamma_k p_{t,B_k}, ~~I_k = [x_k, y_k]
	$$
	Then by Lemma \ref{lem:prefixRelation}, we have $x_k \leq y_k$, $x_{k+1} \leq x_k \leq y_{k+1}$. Therefore, $I_k$ is indeed an interval and $I_k \cap I_{k+1} \neq \emptyset$. Also, $x_{\abs{A}} = f(A, w)$ and $y_1 = 1$. Thus
	$$
	[f(A,w), 1] = \bigcup_{k=1}^{\abs{A}} I_k.
	$$
	Then there exists a $k$ such that $\frac{1}{2}f_{t}^{\ast} \in I_k$:
	$$
	f(B_k,w) \leq \frac{1}{2}f_{t}^{\ast} \leq 1 - \gamma_k + \gamma_k p_{\wedge, t, B_k}
	$$
	Therefore, the results are derived.
\end{proof}

The key difference from the proof of Theorem \ref{thm:main} is the following lemma. This lemma corresponds to Lemma \ref{lem:DeltaEsimateWithP*} and uses $f^{\ast}$ to replace $p^{\ast}$.

\begin{lemma}
	If we have $\gamma_K \geq 1 -  \frac{1}{4} f^{\ast}$, where $f^{\ast} = \min_{1 \leq t \leq T} f_{t}^{\ast}$, then
	\CEqDeltaEstAnd
\end{lemma}
\begin{proof}
	By Lemma~\ref{lem:prefixexists} there exists a prefix $\bB_t$ of $\bA_t$ such that $p_{t, \bB_t} \geq \frac{1}{2}f_{t}^* - 1 + \gamma_{|\bB_t|}$ and $R(t, \bB_t) \geq \frac{1}{2} R(t, \bA_t)$. By the assumption that $1-\gamma_{|\bB_t|} \leq 1-\gamma_K \leq \frac{1}{4} f^{\ast} \leq \frac{1}{4} f^*_{t}  $, we have $p_{t, \bB_t} \geq \frac{1}{4}f_{t}^*$.

	Then, similar to the derivation from Eq.\eqref{eq:addOt} to Eq.\eqref{eq:relaxpstar}, we have
	\begin{align}
		&\EE_t[R(t, \bA_t)] \leq 2 \EE[R(t, \bB_t) \left. \right| \cH_t] \nonumber \\
		=& 2 \EE \left[ \left. R(t, \bB_t) \EE \left[ \left. \frac{1}{p_{t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\} \right| \bB_t \right]  \right| \cH_t \right] \nonumber \\
		=& 2\EE \left[ \left. R(t, \bB_t) \frac{1}{p_{t, \bB_t}} \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right] \nonumber \\
		\leq &\frac{8}{f_{t}^{\ast}} \EE \left[ \left. R(t, \bB_t) \bOne\{\bO_t\geq\abs{\bB_t}\}  \right| \cH_t \right]. \label{eq:btOneotbt}
	\end{align}
	Next, we have
	\begin{equation} \label{eq:andstarbt}
		f(A_t^*, w_t) \leq f(A_t^*,\bU_t) \leq f(\bA_t,\bU_t) \leq f(\bB_t,\bU_t),
	\end{equation}
	where the first inequality is by Lemmas \ref{lem:estimateU} and \ref{lem:andMonotonicity}, the second inequality is by the C$^3$-UCB algorithm, in which the best action $\bA_t$ is selected with respect to $\bU_t$, and the last inequality is by Lemma~\ref{lem:prefixRelation}~(2).

	Then by Lemma \ref{lem:andLip},
	\begin{equation} 
		\label{eq:btuttrue}
		f(\bB_t,\bU_t) \leq f(\bB_t, w_t) + \sum_{k=1}^{\abs{\bB_t}}\gamma_k\bbeta_{t-1}(\delta)\norm{x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}.
	\end{equation}
	Therefore, combining Eq.\eqref{eq:andstarbt} and \eqref{eq:btuttrue}, we have
	\begin{align}
		R(t, \bB_t) & = f_{t}^{\ast} - f(\bB_t, \theta_{\ast}^{\top}x_t) \nonumber \\
		& \leq \sum_{k=1}^{\abs{\bB_t}}\bbeta_{t-1}(\delta)\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}. \label{eq:sumbetax}
	\end{align}
	Finally from Inequalities~\eqref{eq:btOneotbt} and \eqref{eq:sumbetax}， we can derive the result
	\begin{align*}
		&\EE_t[R(t, \bA_t)] \\
		\leq& \frac{8}{f_{t}^*} \EE \left[\bOne\{ \bO_t \geq \abs{\bB_t} \}\sum_{k=1}^{\abs{\bB_t}}\bbeta_{t-1}(\delta)\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right]\\
		\leq& \frac{8}{f^*} \EE \left[ \sum_{k=1}^{ \bO_t } \bbeta_{t-1}(\delta) \norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right].
	\end{align*}
\end{proof}

Then we have the proof of Theorem \ref{thm:and}.

\begin{proof}[of Theorem \ref{thm:and}]
	Similar to the proof of Theorem \ref{thm:main}, we have the following derivation:
	\begin{align}
	&R(n) = \EE \left[ \sum_{t=1}^{n} \EE_{t}[R(t, \bA_t)] \right] \notag\\
	\leq& \EE \left[\sum_{t=1}^{n} \frac{8}{f^{\ast}} \EE_t\left[\bbeta_{t-1}(\delta)\sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right] \right] \label{eq:pfMCdeltaEst}\\
	\leq& \EE\left[\frac{8}{f^{\ast}} \bbeta_{n}(\delta) \sum_{t=1}^{n} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}} \right] \label{eq:pfMCbeta}\\
	\leq& \EE\left[\frac{8}{f^{\ast}} \bbeta_{n}(\delta) \sqrt{ \left(\sum_{t=1}^{n} \bO_t \right) \EE_t \left[\sum_{t=1}^{n} \sum_{k=1}^{\bO_t}\norm{\gamma_k x_{t,\ba_k^t}}_{\bV_{t-1}^{-1}}^2 \right]} \right] \label{eq:pfMCmeaneq}\\
	\leq& \EE \left[\frac{\sqrt{128}}{f^{\ast}} \left(\sqrt{\ln\left(\frac{\det(\bV_n)}{\lambda^d \delta^2}\right) } + \sqrt{\lambda} \right) \right. \notag\\
	&\qquad \qquad \left. \cdot \sqrt{nK \cdot 2\ln \left(\frac{\det(\bV_n)}{\lambda^d} \right)} \right] \label{eq:pfMCsumXi}\\
	\leq& \frac{\sqrt{128}}{f_{\wedge}^{\ast}} \left(\sqrt{d \ln \left( \frac{1 + C_\gamma n/(\lambda d)}{\delta^{2/d}}\right) } + \sqrt{\lambda} \right) \notag\\
	&\qquad \qquad \cdot \sqrt{nKd\ln(1 + C_\gamma n/(\lambda d))}. \label{eq:pfMCdetEst}
	\end{align}
\end{proof}

\end{document}

% This document was modified from the file originally made available by
% Pat Langley and Andrea Danyluk for ICML-2K. This version was
% created by Lise Getoor and Tobias Scheffer, it was slightly modified  
% from the 2010 version by Thorsten Joachims & Johannes Fuernkranz, 
% slightly modified from the 2009 version by Kiri Wagstaff and 
% Sam Roweis's 2008 version, which is slightly modified from 
% Prasad Tadepalli's 2007 version which is a lightly 
% changed version of the previous year's version by Andrew Moore, 
% which was in turn edited from those of Kristian Kersting and 
% Codrina Lauth. Alex Smola contributed to the algorithmic style files.